---
title: "01_MesoMizer_QMS_miniproject"
output: html_document
date: "2023-01-24"
---

```{r}
#install.packages("mizer")
#install.packages("remotes")
#remotes::install_github("sizespectrum/mizerExperimental")
#remotes::install_github("sizespectrum/mizerExperimental", ref = "tuneMR")
library(mizerExperimental)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
```



#Parameter description--------------------------------------------------

```{r}

param_description <- data.frame("parameter" = c("w_inf","w_mat","beta","sigma","R_max","k_vb","l25","l50",
                                                "a","b","catchability","h","k","ks","z0","gamma","w_mat25","erepro"),
                                "description" = c("asymptotic weigth",
                                                  "maturation weight (determines when 50% of the population has matured using a sigmoid function)",
                                                  "preferred predator/prey mass ratio","width of the feeding kernel",
                                                  "Beverton-Holt density dependence parameter",
                                                  "von Bertalanffy growth parameter","length at 25% catch","length at 50% catch",
                                                  "coefficient for age to size conversion",
                                                  "constant for age to size conversion","fisheries efficiency","maximum intake rate","metabolism constant",
                                                  "metabolism coefficient","background mortality coefficient",
                                                  "search volume (obtained from beta and sigma)",
                                                  "weight at which 25% of individuals are mature","reproductive output scalar")
)

knitr::kable(param_description, caption = "parameters' description")
```

## MesoMizer small example - interaction maxtrix based on likelihood of encounter in epipelagic zone only 

```{r}
mesoExample <- data.frame("species" = c("G. braueri", "K. anderssoni", "E. antarctica", "T. macrura", "S. arcticus"),
                           "w_inf" = c(26.6, 13.8, 26.78, 0.318, 0.97), 
                           "k_vb" = c(0.3, 0.7, 0.2, 0.43, 0.8))
                           #"beta" = c(200,200)) #preferred size of prey...small means close to their size
knitr::kable(mesoExample, caption = "species' parameters")

mesoInter <- matrix(c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1), ncol = 5, 
                     dimnames = list(mesoExample$species, mesoExample$species))
knitr::kable(mesoInter, caption = "species' interactions")


```
```{r}
meso_p <- newMultispeciesParams(mesoExample, mesoInter, w_pp_cutoff = 1, )
species_params(meso_p)

mini_proj_params <- species_params(meso_p)

write.csv(mini_proj_params , "mini_project_params.csv")

```


```{r}
meso_p <- steady(meso_p)

plotSpectra(meso_p, power = 1)
plotGrowthCurves(meso_p, species_panel = T)
plotFeedingLevel(meso_p , include_critical = T)
```
```{r}
plotDiet(meso_p,species = "G. braueri")
plotDiet(meso_p,species = "K. anderssoni")
plotDiet(meso_p,species = "E. antarctica")
plotDiet(meso_p,species = "S. arcticus")
plotDiet(meso_p,species = "T. macrura")
```
```{r}
getEReproAndGrowth(meso_p) #calculates the energy rate (g/year) available for reproduction and growth after metabolism and movement
# so we want to know 1- getEReproAndGrowth - wouldn't this just be getMetabolicRate()?
getEGrowth(meso_p) # calculates the energy rate (g/year) available by species and size for growth after metabolism, movement and reproduction have been accounted for
getMetabolicRate(meso_p) #sets the rate at which energy is used for metabolism and activity


plotEnergyBudget(meso_p)
```

#Projection in time----------------------------------------------------------

```{r}

sim_meso <- project(meso_p, t_max = 500)

plotBiomass(sim_meso)
plotDiet(sim_meso@params)

```
#Calibration with empiracle data--------------------------------------------

```{r}
# meso_p@species_params$biomass_observed< <- empirical biomass data per species in g
```


#Setting up Rmax-----------------------------------------------------------

```{r}
meso_p@species_params$R_max <- meso_p@resource_params$kappa * meso_p@species_params$w_inf^-1

meso_p2 <- steady(meso_p)

sim2 <- project(meso_p2, t_max = 500)
plotBiomass(sim2)
meso_p3 <- meso_p2
meso_p3@initial_n <- sim2@n[dim(sim2@n)[1],,] # selecting biomass of all species and sizes at the last time step
```
```{r}
plotDiet(sim2@params)
plotFeedingLevel(sim2)
plotGrowthCurves(sim2, species_panel = T)
```
#Cutting short resource spectrum---------------------------------------------
To change diets: shortcut resources, change interaction matrix, change interaction resource

```{r}
# option 1 doesn't work
resource_params(meso_p3)$w_pp_cutoff <- 1
# option 2 doesn't work
meso_p3 <- mizer::setResource(meso_p3, w_pp_cutoff = 1)
sim3 <- project(meso_p3)
plotBiomass(sim3)
plotSpectra(sim3)
```
#Metabolism loss---------------------------------------------------------------

```{r}
plotEnergyBudget # looking at source code
res <- plotEnergyBudget(meso_p3, return_data = T) # return dataframe for plot
meso_p3@metab[1:3,70:80] #g/year per size bin (amount of abundance per size bin) - rate need to be multiplied by n    (sim@n)
meso_p3@metab[4:5,40:50]

#find metabolic rate (g/year)at each size..
metab_size_f <- meso_p3@metab[1:3,70:100]*sim3@n[1,1:3,70:100]/365 # this should be metabolic loss - g/day of biomass per size....
metab_size_crus <- meso_p3@metab[4:5,40:70]*sim3@n[1,4:5,40:70]/365

#find metabolic loss as % of weight at each size class
#metab_size_f <- meso_p3@metab[1:3,70:100]/meso_p3@w[1:3,70:100] # this should be metabolic loss - g/day of biomass per size....
#metab_size_crus <- meso_p3@metab[4:5,40:70]*sim3@n[1,4:5,40:70]/365

```

#Data wrangle-----------------------------------------------------------------

```{r}

#fish
metab_size_f <- as.data.frame(metab_size_f)
metab_size_f <- metab_size_f %>% 
  rownames_to_column(var = "Species")
metab_size_f <- metab_size_f %>%
  pivot_longer(!Species, names_to = "size", values_to = "metab")

#crustaceans
metab_size_crus <- as.data.frame(metab_size_crus)
metab_size_crus <- metab_size_crus %>% 
  rownames_to_column(var = "Species")
metab_size_crus <- metab_size_crus %>%
  pivot_longer(!Species, names_to = "size", values_to = "metab")



```


#Plot Metabolic loss---------------------------------------------------------

```{r}


ggplot(data=metab_size_f, aes(x=as.numeric(size), y=metab, group=Species)) +
  xlab("Size (g)") +
  ylab("Metabolic loss (g/day)")+
  #scale_x_continuous(breaks = seq(0,2,0.2)) +
  geom_line(aes(color = Species))

ggplot(data=metab_size_crus, aes(x=as.numeric(size), y=metab, group=Species)) +
   xlab("Size (g)") +
  ylab("Metabolic loss (g/day)")+
  scale_x_continuous(breaks = seq(0,2,0.2)) +
  geom_line(aes(color = Species))


```


# Fishing mortality


```{r}
plotFMort(meso_p3) #"knife edge" 
```

```{r}
knitr::kable(gear_params(meso_p3), caption = "gear parameters",row.names = F,digits = 12)
```

```{r}
gear_params(meso_p3)$gear <- c("super_trawler")
gear_params(meso_p3)$knife_edge_size <- 0.5
gear_params(meso_p3)$catchability <- 1
```




```{r}

sim_light <- project(meso_p3, effort = 0.5, t_max = 100)
sim_heavy <- project(meso_p3, effort = 1.5, t_max = 100)

```




```{r}
# assuming sim0 is steady state, sim1 is some kind of variation such as fishing
sim0 = sim_light
sim1 = sim_heavy
fish_sim <- apply(N(sim1)[(dim(N(sim1))[1]-5):(dim(N(sim1))[1]),,],2:3,mean)
unfish_sim <- apply(N(sim0)[(dim(N(sim0))[1]-5):(dim(N(sim0))[1]),,],2:3,mean)

relative_n <- melt((fish_sim - unfish_sim) / (fish_sim + unfish_sim)) 

colnames(relative_n)[1] <- "Species"
legend_levels <- intersect(names(sim0@params@linecolour), relative_n$Species)
ggplot(relative_n) +
  geom_line(aes(x = w, y = value, colour = Species), size = 1) +
  scale_x_continuous(trans = "log10", name = "Weight [g]") +
  scale_y_continuous(name = "Relative difference") +
  scale_color_manual(values = sim0@params@linecolour[legend_levels]) +
  geom_hline(yintercept = 0, linewidth = 1, colour="dark grey", size=0.75)

```

```{r}
plotSpectra2(object1 = sim0, object2 = sim1, name1 =  "light fishing", name2 =  "heavy fishing", 
             total = TRUE, power = 2, ylim = c(1e-8, NA), wlim = c(1e-3, NA))
```



```{r}
biomass<-rbind(getBiomass(sim_light)[100,],getBiomass(sim_heavy)[100,])
row.names(biomass)<-c("light","heavy")
knitr::kable(biomass, caption = "Biomass")
```

#Metabolic loss after fishing--------------------------------------------------


Data wrangle

```{r}

metab_size_f_sim1 <- meso_p3@metab[1:3,70:100]*sim1@n[1,1:3,70:100]/365 # this should be metabolic loss - g/day of biomass per size....
metab_size_crus_sim1 <- meso_p3@metab[4:5,40:70]*sim1@n[1,4:5,40:70]/365

#fish
metab_size_f_sim1 <- as.data.frame(metab_size_f_sim1)
metab_size_f_sim1 <- metab_size_f_sim1 %>% 
  rownames_to_column(var = "Species")
metab_size_f_sim1 <- metab_size_f_sim1%>%
  pivot_longer(!Species, names_to = "size", values_to = "metab2")

#crustaceans
metab_size_crus_sim1 <- as.data.frame(metab_size_crus_sim1)
metab_size_crus_sim1<- metab_size_crus_sim1 %>% 
  rownames_to_column(var = "Species")
metab_size_crus_sim1 <- metab_size_crus_sim1 %>%
  pivot_longer(!Species, names_to = "size", values_to = "metab2")


```



```{r}

ggplot(data=metab_size_f_sim1, aes(x=size, y=metab2, group=Species)) +
  geom_line(aes(color = Species))

ggplot(data=metab_size_crus_sim1, aes(x=size, y=metab2, group=Species)) +
  geom_line(aes(color = Species))

```
#Join dataframes--------------------------------------------------------------

```{r}
all_sim <- metab_size_f %>%
    left_join(metab_size_f_sim1, by='Species') 
```

```{r}

  ggplot(data = all_sim, aes(x = size.x, group = Species)) +
  geom_line(aes(y = metab, group = Species)) +
  geom_line(aes(y = metab2, group = Species))


```





