---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
```


Original code from size structured model to calculate trophic position

Initially setup where dynamic predators and prey are output separately to phytoplankton.

This could be beneficial in mizer, as it will allow the background resource to be treated distinctly and we can assign different baseline values to the background resource.

dim(dynamdiet)
[1]   4 184   4 184

Dimensions of `dynamdiet` are 4 predator groups, 184 predator sizes, 4 prey groups, 184 prey sizes

dim(phytodiet)
[1]   4 184

dimensions of `phytodiet` are 4 predator groups, 184 predator sizes and the value contained is the amount of phytoplankton consumed for each predator and each predator size

```{r}
##### PHYTO DIET DIM1 = SAVE STEP, DIM2 = PRED GROUP, DIM3 = PRED SIZE
##### DYNAM DIET DIM1 = SAVE STEP, DIM2 = PRED GROUP, DIM3 = PRED SIZE, DIM 4 = PREY GROUP, DIM 5 = PREY SIZE

# Average diet array for each predator size class for the final half of the projection
dynamdiet = apply(model$dynam_diet_full[c(floor(0.5 * dim(model$dynam_diet_full)[1]):dim(model$dynam_diet_full)[1]), , , , ], c(2, 3, 4, 5), mean)
phytodiet = apply(model$phyto_diet_full[c(floor(0.5 * dim(model$phyto_diet_full)[1]):dim(model$phyto_diet_full)[1]), , ], c(2, 3), mean)

phyto_tl <- 1 # TL for phytoplankton is 1
start_dynam_tl <- matrix(2, nrow = dim(dynamdiet)[1], ncol = dim(dynamdiet)[2]) # Start TL for dynamic size groups - start at 2 for all groups and sizes

curr_phyto_diet <- phytodiet # Current phyto diet
curr_dynam_diet <- dynamdiet # Current heterotroph diet

total_diet <- apply(curr_phyto_diet, c(1, 2), sum) + apply(curr_dynam_diet, c(1, 2), sum) # Total consumption, in grams wet weight, by pred group and pred size

curr_phyto_frac <- apply(sweep(curr_phyto_diet, c(1, 2), total_diet, '/'), c(1, 2), sum) # Fraction of diet from phyto, by pred group and pred sizes
curr_dynam_frac <- sweep(curr_dynam_diet, c(1, 2), total_diet, '/') # Fraction of diet from each prey group and prey size, for each pred group and pred size

pb = txtProgressBar(
  min = 0,
  max = 100,
  initial = 0,
  style = 3
)

for (j in 1:100) {
  # Gauss-Siedel iterative loop to calculate trophic levels
  setTxtProgressBar(pb, j)
  
  calc_dynam_tl = sweep(curr_dynam_frac, c(3, 4), start_dynam_tl, '*')
  calc_dynam_tl[which(is.nan(calc_dynam_tl) == TRUE)] = 0 # Get rid of nans - these are entries where there is no biomass for a given group
  #calc_dynam_tl[which(calc_dynam_tl == Inf)] = 0 # Get rid of infinite values, occurs with asymptotic size bins, because there is no biomass to have a diet in those bins
  start_dynam_tl = 1 + phyto_tl * curr_phyto_frac + apply(calc_dynam_tl, c(1, 2), sum) # Update trophic level matrix
} # End Gauss-Siedel loop

# Create long form diet dataframe
start_dynam_tl <- melt(start_dynam_tl)

# Assign names to all variables in df_diet
names(start_dynam_tl) <- c("Predator_Species", "Predator_Size_Class", "RTP")
```


Function to get a 4-d diet array from mizer models

```{r}
getDietComp <- function (params,
                         n = initialN(params),
                         n_pp = initialNResource(params),
                         
                         n_other = initialNOther(params),
                         proportion = TRUE)
  
{
  params <- validParams(params)
  pred_kernel(params) <- pred_kernel(params) #required to make this work
  species <- params@species_params$species
  no_sp <- length(species)
  no_w <- length(params@w)
  no_w_full <- length(params@w_full)
  no_other <- length(params@other_encounter)
  other_names <- names(params@other_encounter)
  assertthat::assert_that(identical(dim(n), c(no_sp, no_w)), length(n_pp) ==
                            
                            no_w_full)
  
  n_tot <- sweep(n, 2, params@w * params@dw, "*")
  
  diet_comp <- array(
    0,
    dim = c(dim(n_tot), no_sp + 1, no_w_full),
    
    dimnames = list(
      "Predator" = species,
      
      "wPredator" = params@w,
      
      "Prey" = c(as.character(species), "Resource"),
      #, other_names),
      
      "wPrey" = params@w_full
    )
  )
  
  idx_sp <- (no_w_full - no_w + 1):no_w_full
  
  # if (!is.null(comment(params@pred_kernel))) {
  
  ## code to keep prey size | works only without other backgrounds now
  
  for (iPredator in 1:dim(diet_comp)[1])
    
  {
    for (wPredator in 1:dim(diet_comp)[2])
      
    {
      for (iPrey in 1:no_sp)
        
      {
        for (wPrey in 1:no_w)
          # assuming that w is the tail of w_full, won't work if w_full gets larger than w
          
        {
          diet_comp[iPredator, wPredator, iPrey, wPrey] <- params@pred_kernel[iPredator, wPredator, (idx_sp[wPrey])] *
            
            n_tot[iPrey, wPrey]
          
        }
        
      }
      
    }
    
  }
  
  diet_comp[, , no_sp + 1, ] <- sweep(params@pred_kernel,
                                      
                                      3,
                                      params@dw_full * params@w_full * n_pp,
                                      "*")
  
  inter <- cbind(params@interaction,
                 params@species_params$interaction_resource)
  
  diet_comp[, , 1:(no_sp + 1), ] <- sweep(sweep(diet_comp[, , 1:(no_sp + 1), , drop = FALSE], c(1, 3), inter, "*"),
                                          c(1, 2),
                                          params@search_vol,
                                          "*")
  
  for (i in seq_along(params@other_encounter)) {
    diet_comp[, , no_sp + 1 + i, ] <- do.call(
      params@other_encounter[[i]],
      
      list(
        params = params,
        n = n,
        n_pp = n_pp,
        n_other = n_other,
        
        component = names(params@other_encounter)[[i]]
      )
    )
    
  }
  
  f <- getFeedingLevel(params, n, n_pp)
  
  fish_mask <- n > 0
  
  diet_comp <- sweep(diet_comp, c(1, 2), (1 - f) * fish_mask, "*")
  
  if (proportion) {
    total <- rowSums(diet_comp, dims = 2)
    
    diet_comp <- sweep(diet_comp, c(1, 2), total, "/")
    
    diet_comp[is.nan(diet_comp)] <- 0
    
  }
  
  return(diet_comp)
  
}
```


Load model

```{r}
params_v1 <- readRDS("~/R Projects/MesoMizer/MesoMizer/tuned_params_v3.rds")
```

```{r}
params_v1@w

params_v1@w_full
```



```{r}
plotDiet(params_v1)
```



```{r}
diet_4d <- getDietComp(params_v1)

dim(diet_4d)
```
To adapt for Jon's code, we need to: 
  - add a 'resource' predator to the predator dimension (dimension 1) 
  - might need to update dimension 2 (predator size) to match the full prey size dimension (260 in this case)

```{r}
# diet_frac <- getDietComp(params_v1, proportion = T)

df_diet4d <- melt(diet_4d)

head(df_diet4d)
```

```{r}
# Create a heatmap using ggplot2
df_diet4d %>% 
  group_by(Predator, wPredator, wPrey) %>% 
  summarise(total_consumption = sum(value)) %>% 
  ggplot(aes(x = wPrey, y =  wPredator, fill = total_consumption)) +
  geom_tile() +
  scale_fill_viridis_c() +  # Use a color palette for the heatmap (optional)
  labs(x = "Prey size", y = "Predator size", fill = "value") +
  theme_minimal() +
  facet_wrap(~Predator) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability

glimpse(diet_4d)
```

Since the background resource is now the 14th prey group, we can separate it similarly to phytodiet in the original code:
```{r}

phyto_diet <- diet_4d[, , 14, ]  # Background resource as the 14th prey group

df_phyto_diet <- melt(phyto_diet)

head(df_phyto_diet, 20)

dim(phyto_diet)
```

```{r}
dynam_diet <- diet_4d[, , 1:13, ]  # Dynamic prey excluding background resource

```


Inspect phyto diet

```{r}
dim(phyto_diet)

glimpse(phyto_diet)

summary(phyto_diet)
```
Inspect dynam diet
```{r}
dim(dynam_diet)

glimpse(dynam_diet)

summary(dynam_diet)
```

```{r}
curr_phyto_diet <- phyto_diet
curr_dynam_diet <- dynam_diet
```



```{r}
total_diet <- apply(curr_phyto_diet, c(1,2), sum) + apply(curr_dynam_diet, c(1,2), sum)
```


```{r}
curr_phyto_frac <- apply(sweep(curr_phyto_diet, c(1,2), total_diet, '/'), c(1,2), sum) # Fraction of diet from phyto, by pred group and pred sizes
curr_dynam_frac <- sweep(curr_dynam_diet, c(1,2), total_diet, '/') # Fraction of diet from each prey group and prey size, for each pred group and pred size
```


```{r}
# Create a new array with 260 predator sizes, padding the first 160 sizes with zeros
new_phyto_diet <- matrix(0, nrow = dim(curr_phyto_frac)[1], ncol = 260)

# Place the summed background resource values into the last 100 positions of predator sizes
new_phyto_diet[, 161:260] <- curr_phyto_frac

# Result: new_phyto_diet has dimensions (13 predator groups, 260 predator sizes)

```




Trophic Level Calculation (Gauss-Seidel Iteration)
Use the Gauss-Seidel method to calculate the trophic levels iteratively. The initial start_dynam_tl matrix can still begin at a trophic level of 2:

```{r}
phyto_tl <- 2.3 # TL for background resource, informed by POM SIA results

start_dynam_tl <- matrix(2, nrow = dim(curr_dynam_frac)[1], ncol = dim(curr_dynam_frac)[2]) # Start TL for dynamic size groups - start at 2 for all groups and sizes

# pb = txtProgressBar(min = 0, max = 100, initial = 0, style = 3)
  
  for(j in 1:100){ # Gauss-Siedel iterative loop to calculate trophic levels
    setTxtProgressBar(pb, j)
    
    calc_dynam_tl = sweep(curr_dynam_frac, c(3,4), start_dynam_tl, '*')
    calc_dynam_tl[which(is.nan(calc_dynam_tl) == TRUE)] = 0 # Get rid of nans - these are entrys where there is no biomass for a given group
    calc_dynam_tl[which(calc_dynam_tl == Inf)] = 0 # Get rid of infinite values, occurs with asymptotic size bins, because there is no biomass to have a diet in those bins
    start_dynam_tl = 1+phyto_tl*curr_phyto_frac + apply(calc_dynam_tl, c(1,2), sum) # Update trophic level matrix
  } # End Gauss-Siedel loop
```

Convert to Long-Form Dataframe
Finally, convert start_dynam_tl to long-form, which will now reflect the 13 predator groups and 100 predator sizes:

```{r}
start_dynam_tl_df <- melt(start_dynam_tl)
names(start_dynam_tl_df) <- c("Predator_Species", "Predator_Size_Class", "RTP")

summary(start_dynam_tl_df)
```


```{r}

# make predators and prey factor variables
start_dynam_tl_df$Predator_Species <- as.factor(start_dynam_tl_df$Predator_Species)

df_size_classes <- params_v1@dw # Create new object with list of size classes in log10 g
df_size_classes <- as.data.frame(df_size_classes) # Convert to a dataframe
names(df_size_classes) <- c("Predator_Size") # Create new name Size for the variable
df_size_classes$Predator_Size_Class <- 1:length(params_v1@dw)

df_species_predator <- params_v1@species_params$species # Create new object with species names from model
df_species_predator <- as.data.frame((df_species_predator)) # Convert to a dataframe
names(df_species_predator) <- c("Predator") # Rename variable
df_species_predator$Predator_Species <- 1:length(params_v1@species_params$species) # Create new integer variable in ascending order with 'Species'
df_species_predator$Predator_Species <- as.factor(df_species_predator$Predator_Species) # Convert to factor
df_species_predator$Predator <- as.factor(df_species_predator$Predator) # Convert to factor

glimpse(df_size_classes)
glimpse(df_species_predator)

# ## join df_N with df_size_class data in order to create log10 size classes rather then the size     class integer 1:184 that is in df_N
# df_combined <- left_join(start_dynam_tl_df, df_size_classes, by = "Predator_Size_Class") %>%
#   
#   left_join(df_species_predator, by = "Predator_Species") 

## join df_N with df_size_class data in order to create log10 size classes rather then the size     class integer 1:184 that is in df_N
df_combined <- left_join(start_dynam_tl_df, df_species_predator, by = "Predator_Species") 
```



```{r}
ggplot(data = start_dynam_tl_df, aes(x = Predator_Size_Class, y = RTP, colour = Predator_Species)) +
  geom_point() +
  theme_bw() +
  facet_wrap( ~ Predator_Species, scales = "free") +
  theme(
    legend.position = "none"
  )
```




























