---
title: "05_MesoMizer_Shelf_ther"
output: html_document
date: "2024-08-13"
---

This script contains MizerShelf and therMizer packages.


```{r}
library(mizerExperimental)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(reshape2)
# remotes::install_github("gustavdelius/mizerShelf")
library(mizerShelf)
#install.packages("remotes")
#remotes::install_github("sizespectrum/therMizer")
library(therMizer)
#install.packages("sensobol")
library(sensobol)
library(sensitivity)
library(dietr)
library(rgl)
library(viridis)
library(patchwork)
library(scales)
library(stringr)

 
 rm(list =ls()) # clear memory
 
Biomass_size <- read.csv("Biomass_persize.csv") 
```

#Parameter description--------------------------------------------------

```{r}

param_description <- data.frame("parameter" = c("w_inf","w_mat","beta","sigma","R_max","k_vb","l25","l50",
                                                "a","b","catchability","h","k","ks","z0","gamma","w_mat25","erepro", "RQ", "gear_mort", "p_FC", "z0", "rho"),
                                "description" = c("asymptotic weigth",
                                                  "maturation weight (determines when 50% of the population has matured using a sigmoid function)",
                                                  "preferred predator/prey mass ratio","width of the feeding kernel",
                                                  "Beverton-Holt density dependence parameter",
                                                  "von Bertalanffy growth parameter","length at 25% catch","length at 50% catch",
                                                  "coefficient for age to size conversion",
                                                  "constant for age to size conversion","fisheries efficiency","maximum intake rate","metabolism constant",
                                                  "metabolism coefficient","background mortality coefficient",
                                                  "search volume (obtained from beta and sigma)",
                                                  "weight at which 25% of individuals are mature","reproductive output scalar", "respiratory quotient", "mortality rate due to fishing gear", "carbon content of fecal pellets (%)", "carrion parameter", "carrion parameter")
) #add biomass observed 

knitr::kable(param_description, caption = "parameters' description")


```

#Species parameters


```{r}
sp_1 <-
  data.frame(
    "species" = c("Euphausia similis","Thysannoessa macura","Sergestes spp", "Themisto spp", "Lampanyctus australis","Nannobrachium achirus","Cyclothone spp","Bathylagus spp","Cranchidae","Chaetognath","Gymnosomata","Pyrosoma atlanticum","Atolla wyvillei"),
    "w_inf" = c(0.417,0.318,1.7,0.1,26.78,58.3,3.0,81.43,20,1.0,0.28,375.8,12), #check cyclothone in fishbase & chaetognath in sealifebase 
    "beta" = c(15849000, 15849000, 100, 100, 100, 100, 100, 100, 25, 22, 22, 50119000, 5012), #lower ppmr for piscivores, higher for omnivores and planktivores
    "biomass_observed" = c(0.1970, 0.1582, 0.0704,0.1012,0.6665,0.0658,0.4307,0.1493,0.2333,4.3799,0.0126,31.4392,0.1215), #biomass in g m-2. See Biomass_SOLACE.rmd
    "k_vb" = c(0.6, 0.4, 0.8, 0.6, 0.2, 0.4, 0.9, 0.5, 0.4, 0.2, 0.2,0.2,0.2),
    "gear_mort" = c(0,0,0,0,0,0,0,0,0,0,0,0,0),
    "RQ" = c(1.7, 1.7, 1.45, 1.45, 2.5, 2.5, 2.5, 2.5, 2.6, 1.67, 1.58, 2.8, 0.97), #see carbon spreadsheet for references
    #"p_FC" = c(0.4, 0.4, 0.4,0.4, 0.4, 0.4,0.4, 0.4, 0.4,0.4, 0.4, 0.4,0.4), # carbon content of fecal pellets (%)
    "p_FC" = c(0.2, 0.2, 0.2,0.2, 0.2, 0.2,0.2, 0.2, 0.2,0.2, 0.2, 0.2,0.2), # carbon content of fecal pellets (%)
    "z0" = c(0.1, 0.1, 0.1,0.1, 0.1, 0.1, 0.1, 0.1, 0.1,0.1, 0.1, 0.1, 0.1), # carrion parameter - may delete
   "rho" = c(200, 200, 200, 200, 200, 200,200, 200, 200,200, 200, 200,200)) # carrion parameter - may delete
#"sigma" = c(0.7,0.7, 1.3,1.3,1.3,1.3,1.3,1.3,1.3,0.46,0.57,0.7,0.52)) #- this causes p. atlanticum to go extinct(too narrow... factor these in later)

              
knitr::kable(sp_1, caption = "species' parameters")


                
                   
```


# Species interaction matrix

### add in DVM based on % of community at each depth
```{r}
#predator = row, prey = column
#sp_2_inter reflects feeding type and DVM
sp_2_inter <-matrix <- matrix(c(
  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
  0.6700000000000002, 0.55, 1, 0.8900000000000001, 0, 0, 0, 0, 0, 0.66, 0.35, 0.65, 0.4,
  0.78, 0.6599999999999999, 0.8900000000000001, 1, 0, 0, 0, 0, 0, 0.7699999999999999, 0.29, 0.6000000000000001, 0.35,
  0.42, 0.30000000000000004, 0.62, 0.57, 0.9999999999999999, 0.22, 0.304, 0.22, 0.45, 0.41000000000000003, 0.22, 0.78, 0.7799999999999999,
  0.07, 0.03, 0.35, 0.29, 0.22, 1, 0.916, 1, 0.2, 0.13, 1, 0.03, 0.21,
  0.15400000000000003, 0.114, 0.434, 0.374, 0.304, 0.916, 1, 0.916, 0.28400000000000003, 0.21400000000000002, 0.916, 0.114, 0.238,
  0.07, 0.03, 0.35, 0.29, 0.22, 1, 0.916, 1, 0.2, 0.13, 1, 0.03, 0.21,
  0.8700000000000001, 0.8300000000000001, 0.7, 0.81, 0.45, 0.2, 0.28400000000000003, 0.2, 1, 0.93, 0.2, 0.5, 0.24000000000000002,
  0.9300000000000002, 0.8900000000000001, 0.66, 0.7699999999999999, 0.41000000000000003, 0.13, 0.21400000000000002, 0.13, 0.93, 1, 0.13, 0.53, 0.2,
  0.07, 0.03, 0.35, 0.29, 0.22, 1, 0.916, 1, 0.2, 0.13, 1, 0.03, 0.21,
  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
  0.21000000000000002, 0.09, 0.4, 0.35, 0.7799999999999999, 0.21, 0.238, 0.21, 0.24000000000000002, 0.2, 0.21, 0.5700000000000001, 1
), nrow = 13, byrow = TRUE,
  dimnames = list(sp_1$species, sp_1$species) )

knitr::kable(sp_2_inter, caption = "species' interactions")

# Convert the interaction matrix to a data frame
sp_2_inter_df <- as.data.frame(sp_2_inter)

# Ensure row names are preserved as a column (optional)
sp_2_inter_df$Predator <- rownames(sp_2_inter_df)

# Rearrange columns so 'Predator' is the first column (optional)
#sp_2_inter_df <- sp_2_inter_df[, c(ncol(sp_2_inter_df), 1:(ncol(sp_2_inter_df) - 1))]

write.csv(sp_2_inter_df, "sp_2_interactions.csv", row.names = FALSE)

# Print the data frame using knitr::kable for display
knitr::kable(sp_2_inter_df, caption = "Species' interactions")


```



# Tune params

```{r}
 
 # sp_2 <- newMultispeciesParams(sp_1, sp_2_inter, w_pp_cutoff = 0.01)
 # params_v1 <- setParams(sp_2)
 # params_v1 <- tuneParams(sp_2)

params_v0 <- readRDS("tuned_params_DVM3.rds")

# Update the "p_FC" parameter

#"Euphausia similis","Thysannoessa macura","Sergestes spp", "Themisto spp", "Lampanyctus australis","Nannobrachium achirus","Cyclothone spp","Bathylagus #spp","Cranchidae","Chaetognath","Gymnosomata","Pyrosoma atlanticum","Atolla wyvillei"

params_v0@species_params$p_FC <- c(0.15, 0.15, 0.15, 0.15, 0.07, 0.07, 0.07, 0.07, 0.07, 0.45, 0.07, 0.22, 0.07) #revisit this.... 
params_v0@species_params$RQ <- c(1.29, 1.29, 1.4, 0.97, 0.9, 0.9, 0.9, 0.9, 0.97, 0.97, 0.8, 1.16, 0.97) #see carbon spreadsheet for references

# Save the updated object
saveRDS(params_v0, "tuned_params_DVM3_updated.rds")

# Check the change
params_v0@species_params$p_FC
params_v0@species_params$RQ

params_v0 <- tuneParams(params_v0)


```




# Thermizer -> set thermal tolerances for species

```{r}

# Assign them thermal tolerance limits - general thresholds - will look these up
temp_min <- c(-5, -5,-5, -5,-5, -5,-5, -5,-5, -5,-5, 0,-5)
temp_max <- c(13, 13,13, 13,13, 13,13, 13,13, 12,13, 20,13) #revisit
species_params(params_v0)$temp_min <- temp_min
species_params(params_v0)$temp_max <- temp_max
```


### set temperature parameters for DVM

```{r}
realm_names <- c("Epi","Umeso","Lmeso")
species_names <- as.character(params_v0@species_params$species)
sizes <- params_v0@w

# Create the vertical migration array and fill it
vertical_migration_array <- array(0, dim = (c(length(realm_names), 
                                  length(species_names), length(sizes))), 
                                  dimnames = list(realm = realm_names, sp = species_names, 
                                  w = signif(sizes,3))) # realm x species x size

Epi <- which(realm_names == "Epi")
Umeso <- which(realm_names == "Umeso") 
Lmeso <- which(realm_names == "Lmeso") 

# Define a list of species and their corresponding time allocations - based on average % per depth from SOLACE (Biomass_SOLACE.rmd)
species_allocations <- list(
  "Lampanyctus australis" = c(Epi = 0.21, Umeso = 0.57, Lmeso = 0.22),
  "Euphausia similis" = c(Epi = 0.79, Umeso = 0.14, Lmeso = 0.07),
  "Thysannoessa macura" = c(Epi = 0.91, Umeso = 0.06, Lmeso = 0.03),
  "Sergestes spp" = c(Epi = 0.46, Umeso = 0.19, Lmeso = 0.35),
  "Themisto spp" = c(Epi = 0.57, Umeso = 0.14, Lmeso = 0.29),
  "Nannobrachium achirus" = c(Epi = 0, Umeso = 0, Lmeso = 1),
  "Cyclothone spp" = c(Epi = 0.056, Umeso = 0.028, Lmeso = 0.916), 
  "Bathylagus spp" = c(Epi = 0, Umeso = 0, Lmeso = 1),
  "Cranchidae" = c(Epi = 0.76, Umeso = 0.04, Lmeso = 0.20),
  "Chaetognath" = c(Epi = 0.80, Umeso = 0.07, Lmeso = 0.13),
  "Gymnosomata" = c(Epi = 0, Umeso = 0, Lmeso = 1),
  "Pyrosoma atlanticum" = c(Epi = 0.43, Umeso = 0.54, Lmeso = 0.03),
  "Atolla wyvillei" = c(Epi = 0, Umeso = 0.79, Lmeso = 0.21)
)

# Loop through the species and populate the vertical migration array
for (sp in names(species_allocations)) {
  sp_index <- which(species_names == sp)
  allocations <- species_allocations[[sp]]
  vertical_migration_array[Epi, sp_index, ] <- allocations["Epi"]
  vertical_migration_array[Umeso, sp_index, ] <- allocations["Umeso"]
  vertical_migration_array[Lmeso, sp_index, ] <- allocations["Lmeso"]
}

```


### set up exposure array

```{r}
exposure_array <- array(0, dim = (c(length(realm_names), length(species_names))), 
                  dimnames = list(realm = realm_names, sp = species_names)) # realm x species

for (r in seq(1,length(realm_names),1)) {
    for (s in seq(1,length(species_names),1)) {
        if (any(vertical_migration_array[r,s,] > 0)) {
            exposure_array[r,s] = 1
        }
    }
}


# Create temperature array with constant values per depth layer
times <- 0:500
ocean_temp_array <- array(NA, dim = c(length(times), length(realm_names)), 
                          dimnames = list(time = times, realm = realm_names))

# Set constant temperatures for each realm
for (i in 1:length(times)) {
  ocean_temp_array[i,] <- c(10.5, 10, 7.5) # 10.5°C for Epi, 10°C for Umeso, 7.5°C for Lmeso
}

```


### upgrade therParams - (incorporate temp)

```{r}


params_v1 <- upgradeTherParams(params = params_v0, 
                            temp_min = temp_min,
                            temp_max = temp_max,
                            ocean_temp_array = ocean_temp_array,
                           # n_pp_array = n_pp_array, 
                            vertical_migration_array = vertical_migration_array,
                            exposure_array = exposure_array, 
                            aerobic_effect = TRUE, 
                            metabolism_effect = TRUE)
                                

species_params_df <- params_v1@species_params
write.csv(species_params_df, file = "species_params.csv", row.names = FALSE)


```


#Supplementary plots---------------------------------------------------------------------

```{r}

sim_2 <- project(params_v1)




plotSpectra(params_v1, power = 0, total = FALSE) + 
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),  # Adjust size of axis text
    axis.title = element_text(size = 16), # Adjust size of axis titles
    legend.text = element_text(size = 12), # Adjust size of legend text
    legend.title = element_text(size = 14) # Adjust size of legend title
  )


ggsave("Supp_fig1_ss0.png")


plotSpectra(params_v1, power = 1, total = TRUE) + 
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),  # Adjust size of axis text
    axis.title = element_text(size = 16), # Adjust size of axis titles
    legend.text = element_text(size = 12), # Adjust size of legend text
    legend.title = element_text(size = 14) # Adjust size of legend title
  )


ggsave("Supp_fig2_ss1.png")


plotSpectra(params_v1, power = 2, total = TRUE) + theme_minimal() #slope should be close to zero - biomass constant over log size bins

ggsave("Supp_fig3_ss2.jpeg")

#plotDiet(params_v1)



# Extract diet proportions
diet_plot <- plotDiet(params_v1, return_data = TRUE)

# Modify the plot to allow free x-axis scaling
ggplot(diet_plot, aes(x = w, y = Proportion, fill = Prey)) +
  geom_area() +
  facet_wrap(~ Predator, scales = "free_x") +  # Ensure unique x-axis per species
  scale_x_log10() +  # Keep log scale
  theme_minimal() +
  labs(x = "Size [g]", y = "Proportion")




ggsave("Supp_fig4_diet.png")

plotDeath(params_v1, proportion = T)   + theme_minimal()
ggsave("Supp_fig5_death.png")


plotBiomassObservedVsModel(params_v1, labels = FALSE) + theme_minimal()
ggsave("Supp_fig6_biomassobservedvmodel.png")


plotGrowthCurves(params_v1, species_panel = T)  + theme_minimal()
ggsave("Supp_fig7_growth_curves.png")

plotEnergyBudget(params_v1)

slope_biomass <- getCommunitySlope(sim_2)

slope_biomass[1,]
slope_biomass[100,]

plot(slope_biomass)

```
#validate model with size data from observations
## 1. convert length to weight

```{r}
# from Ezekiel_ESS code

# Step 1: Extract lower and upper bounds of each bin

Biomass_size <- Biomass_size %>%
  mutate(
    bin = str_remove_all(bin, "\\[|\\]|\\(|\\)")  # remove brackets and parentheses
  ) %>%
  separate(bin, into = c("L_bin_min", "L_bin_max"), sep = ",", convert = TRUE) #length in cm

# Step 1: Create a new dataframe with only necessary species parameters
L_W_df <- species_params_df %>%
  select(species, a, b)

#update a and b
L_W_df <- L_W_df %>%
  mutate(
    a = if_else(species == "Nannobrancium achirus", 0.00451, a),
    b = if_else(species == "Nannobrancium achirus", 3, b)
  )


# Step 2: Join species parameters to your Biomass_size data
Biomass_size <- Biomass_size %>%
  left_join(L_W_df, by = c("Taxa_updated" = "species"))


# Step 3: Calculate min and max weight per bin
Biomass_size <- Biomass_size %>%
  mutate(
    weight_min_obs = round(a * L_bin_min^b, 4),
    weight_max_obs = round(a * L_bin_max^b, 4))
  # ) %>%
  # select(Taxa_updated, L_bin_min, L_bin_max, count, proportion, abd_total_gm2_sp, abundance_per_bin, biomass_observed, a, b, weight_min_obs,weight_max_obs )

Biomass_size <- Biomass_size %>%
  mutate(
    # Step 1: Midpoint of each length bin
    weight_mid = (weight_min_obs + weight_max_obs) / 2
  ) 


ggplot(Biomass_size, aes(x = weight_mid, y = biomass_per_bin_g)) +
  geom_col() +
  scale_x_log10() +  # optional: log scale for weight
  scale_y_continuous() +
  facet_wrap(~ Taxa_updated, scales = "free_y") +
  labs(
    x = "Weight (g, mid bin)",
    y = "Biomass per bin (g/m²)",
    title = "Size frequency distribution by taxon"
  ) +
  theme_bw()

```
#2. compare to modeled size spectra ----- stopped here - need different a and b from default

```{r}



# STEP 1: Define log10 bin edges, midpoints, and widths
breaks <- 10^seq(-3, 3, 0.1)  # log-scaled weight breaks from 0.001 to 1000 g
midpoints <- (head(breaks, -1) + tail(breaks, -1)) / 2
widths <- diff(breaks)
labels <- cut(midpoints, breaks = breaks, include.lowest = TRUE)

# STEP 2: Create a bin lookup table
bin_lookup <- data.frame(
  weightclass = as.character(labels),
  bin_midpoint = midpoints,
  bin_width = widths
)

# STEP 3: Bin observed weights into these log-scale bins
Biomass_size <- Biomass_size %>%
  mutate(
    weightclass = cut(weight_mid, breaks = breaks, include.lowest = TRUE)
  )

# STEP 4: Join bin midpoints and widths into your observed data
Biomass_size <- Biomass_size %>%
  mutate(weightclass = as.character(weightclass)) %>%
  left_join(bin_lookup, by = "weightclass")

# STEP 5: Normalize biomass per bin by bin width
Biomass_size <- Biomass_size %>%
  mutate(
    norm_biomass = biomass_per_bin_g / bin_width
  )

# STEP 6: Rename species column to match Mizer
Biomass_size <- Biomass_size %>%
  rename(Species = Taxa_updated)

# STEP 7: Summarize biomass per bin per species
mean_biomass_size_spectra <- Biomass_size %>%
  group_by(bin_midpoint, Species) %>%
  summarise(mean_nbiomass = mean(norm_biomass, na.rm = TRUE)) %>%
  ungroup()

# STEP 8: Plot observed vs. modeled size spectra
plotSpectra(params_v1, power = 0, resource = FALSE) +
  geom_point(data = mean_biomass_size_spectra,
             aes(x = bin_midpoint, y = mean_nbiomass, colour = Species)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~ Species, scales = "free_y") +
  theme_bw() +
  labs(x = "Weight (g)", y = "Normalized Biomass (g/m² per bin width)")

```

```{r}
# STEP 8: Plot observed vs. modeled size spectra (combined plot)
plotSpectra(params_v1, power = 4, resource = FALSE) +
  geom_point(data = mean_biomass_size_spectra,
             aes(x = bin_midpoint, y = mean_nbiomass, colour = Species)) +
  scale_x_log10() +
  scale_y_log10() +
  theme_bw() +
  labs(x = "Weight (g)", y = "Normalized Biomass (g/m² per bin width)")

```


#fraction of biomass thats a function of size
```{r}
# 
# # 1. Run your Mizer model
# results <- project(params_v1)
# 
# # 2. Extract final timestep biomass
# n_final <- results@n[nrow(results@n), , ]
# sizes <- params_v1@w
# biomass <- n_final * sizes
# 
# # 3. Total community biomass at each size
# total_biomass_size <- colSums(biomass)
# total_biomass_size[total_biomass_size == 0] <- NA
# 
# # 4. Calculate fraction of total community biomass per species and size
# fraction_matrix <- biomass / matrix(rep(total_biomass_size, each = nrow(biomass)),
#                                     nrow = nrow(biomass))
# 
# # 5. Define fine log-size grid
# size_grid <- 10^seq(log10(min(sizes)), log10(max(sizes)), length.out = 300)
# 
# # 6. Lognormal fit function with NO filtering
# fit_lognormal_fraction <- function(size, frac_vec, grid) {
#   if (sum(frac_vec, na.rm = TRUE) == 0) return(rep(NA, length(grid)))  # skip species with zero biomass
#   
#   log_size <- log10(size)
#   log_mean <- weighted.mean(log_size, frac_vec, na.rm = TRUE)
#   log_sd <- sqrt(weighted.mean((log_size - log_mean)^2, frac_vec, na.rm = TRUE))
#   
#   # handle edge cases
#   if (is.na(log_mean) || is.na(log_sd) || log_sd == 0) return(rep(NA, length(grid)))
#   
#   peak_contribution <- sum(frac_vec, na.rm = TRUE)
#   d <- dnorm(log10(grid), mean = log_mean, sd = log_sd)
#   (d / sum(d, na.rm = TRUE)) * peak_contribution  # scaled to real contribution
# }
# 
# # 7. Apply smoother to ALL species (no filter)
# fraction_fits <- lapply(1:nrow(fraction_matrix), function(i) {
#   fit_lognormal_fraction(sizes, fraction_matrix[i, ], size_grid)
# })
# names(fraction_fits) <- rownames(fraction_matrix)
# 
# # 8. Keep only species with actual biomass (not all-NA)
# valid <- sapply(fraction_fits, function(x) !all(is.na(x)))
# fraction_matrix_smooth <- do.call(cbind, fraction_fits[valid])
# species_names <- names(fraction_fits)[valid]
# 
# # 9. Color palette for species (needed!)
# library(viridis)
# colors <- viridis(ncol(fraction_matrix_smooth), option = "D")
# 
# # 10. Save as JPEG (since matplot is base graphics)
# jpeg("biomass_size_fraction.jpeg", width = 1200, height = 800, res = 150)
# 
# # Increase right margin for external legend
# par(mar = c(5, 4, 4, 12), xpd = TRUE)
# 
# # Plot
# matplot(size_grid, fraction_matrix_smooth, type = "l", log = "x",
#         lty = 1, lwd = 2,
#         col = colors,
#         xlab = "Size (g)", ylab = "Fraction of Total Community Biomass",
#         main = "")
# 
# legend("topright", inset = c(-0.5, 0), legend = species_names,
#        col = colors, lty = 1, lwd = 2, cex = 0.75, bty = "n")
# 
# dev.off()
# 
# png("biomass_size_fraction.png", width = 1200, height = 800, res = 150)

```




# Carbon outputs

### 1. Respiratory carbon (DIC)

#### 1a. Get metabolic rates per species & size

```{r}
#Get metabolic rate per species and size (g m-2 y-1)
metabolic_rates <- metab(params_v1)

# Convert metabolic rates array to long-format data frame
metabolic_rates_df <- melt(metabolic_rates, varnames = c("Species", "Size"), value.name = "Metabolic_Rate")

# Plot with log scale for size
ggplot(metabolic_rates_df, aes(x = Size, y = Metabolic_Rate, color = as.factor(Species))) +
  geom_line() +
  scale_x_log10() +
  labs(title = "Metabolic Rate per Size for Each Species",
       x = "Size (w)",
       y = "Metabolic Rate",
       color = "Species") +
  theme_minimal()
```

### 1.b incorporate abundances (n_initial)

```{r}
# Function to calculate respired carbon per species and size
respired_carbon_per_species <- function(params_v1, 
                                        n = params_v1@initial_n, 
                                        n_other = params_v1@initial_n_other) {
    # Get metabolic rates per species and size (g m-2 y-1)
    metabolic_rates <- metab(params_v1)
    
    # Multiply metabolic rates by abundance for each species and size
    respired_carbon <- sweep(metabolic_rates * n, 2, params_v1@dw, "*", check.margin = FALSE)
    
    # Return the respired carbon per species and size
    return(respired_carbon)
}

# Call the function to get respired carbon
respired_carbon <- respired_carbon_per_species(params_v1)

species_names <- rownames(params_v1@species_params)
size_bins <- params_v1@w

# Extract the respiration matrix
#respiration_matrix <- respired_carbon$respired_carbon

# Assign the correct column names (size bins)
colnames(respired_carbon) <- size_bins

# Convert to a data frame and add species names
respiration_df <- as.data.frame(respired_carbon)
respiration_df$Species <- rownames(respired_carbon)

# Reshape the data to long format

respiration_long <- respiration_df %>%
  pivot_longer(cols = -Species, names_to = "Size", values_to = "respiration_production") %>%
  mutate(Size = as.numeric(Size))

# View the final data frame
print(respiration_long)

# Plot using ggplot2 with log scale for size
RC_p <- ggplot(respiration_long, aes(x = Size, y = respiration_production, color = Species)) +
  geom_line() +
  scale_x_log10() +
  labs(title = "Respired weight (g) per Size for Each Species",
       x = "Size (g)",
       y = "g C m-2 h-1",
       color = "Species") +
  theme_minimal()


```


## 1c. Convert Mizer metabolic rate (g y-1) to respiratory oxygen uptake (L h-1) then respired carbon (g C m-2 h-1)


```{r}

#RC = RO * RQ * (12/22.4) #Kwong et al. 2020
    #RC: respiratory carbon (g C ind-1 h-1); 
    #RO: respiratory oxygen uptake (L O2 ind-1 h-1); 
    #RQ: respiratory quotient (unit-less);
    #12: molar weight of carbon (g mol-1); 
    #22.4: molar volume of an ideal gas at standard pressure (mol L-1)
    #TD = time at depth

#to convert metabolic rate (respiration_production) to RO - must convert to energy (KJ) using these assumptions:

# Constants ----- include reference
energy_content <- 23.7         # kJ per gram based on a mixed diet - could make this more dynamic to reflect feeding types
oxygen_conversion <- 19.4      # kJ per liter of O2
hours_per_year <- 8760         # Number of hours in a year


respired_carbon_df <- respiration_long %>% 
  mutate(oxygen_uptake_liters_h = respiration_production * energy_content/oxygen_conversion/hours_per_year)

# Define the conversion factor from oxygen uptake to carbon
oxygen_to_carbon <- 12 / 22.4  # This is the factor (12/22.4) to convert oxygen to carbon

# Add RQ values to the respired_carbon_df
respired_carbon_df$RQ <- params_v1@species_params$RQ[match(respired_carbon_df$Species, params_v1@species_params$species)]

# View the updated dataframe
head(respired_carbon_df)


respired_carbon_df <- respired_carbon_df %>% 
  mutate(respired_carbon = oxygen_uptake_liters_h * RQ * oxygen_to_carbon)




```



#### plot

```{r}
# Plot using ggplot2 with log scale for size
RC_p <- ggplot(respired_carbon_df, aes(x = Size, y = respired_carbon, color = Species)) +
  geom_line() +
  scale_x_log10() +
  labs(title = "Respired Carbon per Size for Each Species",
       x = "Size (g)",
       y = "g C m-2 h-1",
       color = "Species") +
  theme_minimal()

# Display the plot
print(RC_p)

#ggsave("RC.png", plot = RC_p)

```


# 2. Fecal Carbon (POC)

## 2a. Extract feces production per species and size

```{r}

detritus_production_per_species <- function(params_v1, n = params_v1@initial_n, # in MizerShelf, detritus production includes feces, decomposing carrion and external. For now we only want to extract feces, but we'll keep the other functions in at this stage.
                                            n_other = params_v1@initial_n_other, #resource?
                                            rates = getRates(params_v1)) {
    # Calculate the consumption for each species
    consumption <- sweep((1 - rates$feeding_level) * rates$encounter * n, 2,
                         params_v1@dw, "*", check.margin = FALSE)
    
    # Calculate the feces production for each species
    feces <- sweep(consumption, 1, (1 - params_v1@species_params$alpha), "*", 
                   check.margin = FALSE)
    
    # Calculate the carrion decomposition
    carrion <- params_v1@other_params$carrion$decompose * n_other$carrion
    
    # Create a list to store detritus production per species
    detritus_per_species <- list(
        external = params_v1@other_params$detritus$external,
        feces = feces,
        carrion = carrion
    )
    
    return(detritus_per_species)
}



detritus <- detritus_production_per_species(params_v1)

species_names <- rownames(params_v1@species_params)
size_bins <- params_v1@w

# Extract the feces matrix
feces_matrix <- detritus$feces

# Assign the correct column names (size bins)
colnames(feces_matrix) <- size_bins

# Convert to a data frame and add species names
feces_df <- as.data.frame(feces_matrix)
feces_df$Species <- rownames(feces_matrix)

# Reshape the data to long format

feces_long <- feces_df %>%
  pivot_longer(cols = -Species, names_to = "Size", values_to = "Feces_production") %>%
  mutate(Size = as.numeric(Size))

# View the final data frame
print(feces_long)



```



## 2b. Convert feces (g m-2 y-1) to (g m-2 C h-1)


```{r}

# Extract the p_FC (carbon % per fecal pellet) parameter from params_v1
p_FC <- params_v1@species_params$p_FC

# Create a data frame with species and their corresponding p_FC values
p_FC_df <- data.frame(Species = rownames(params_v1@species_params), p_FC = p_FC)

# Merge the p_FC values into the feces_long dataframe
feces_long <- merge(feces_long, p_FC_df, by = "Species", all.x = TRUE)


# Calculate Fecal Carbon (FC)
feces_df <- feces_long %>%
  mutate(Fecal_Carbon = Feces_production *p_FC/hours_per_year,
          Fecal_Carbon_mg = Fecal_Carbon * 1000)  # Convert to milligrams)  


feces_p <- ggplot(feces_df, aes(x = Size, y = Fecal_Carbon_mg, color = as.factor(Species))) +
  geom_line() +
  scale_y_log10()+
  scale_x_log10() +
  labs(title = "Fecal Carbon per Size for Each Species",
       x = "Size (g)",
       y = "mg C m-2 h-1",
       color = "Species") +
  theme_minimal()

feces_p

#ggsave("FC.png", plot = feces_p,  width = 7, height = 5)

```



# 3. Carrion carbon (POC)

## 3a. Approach 1: take carrion from proportion of detritus. However, this includes a degradation term, so we may miss "fresh" carrion. Also, if used with degradation term we may not need to apply remineralization rates later... i.e. this "POC" is actually "DIC" and remains suspended.

```{r}
#approach 1 - take carrion from proportion of detritus

# # Extract the feces matrix
# carrion_matrix <- detritus$carrion
# 
# # Assign the correct column names (size bins)
# colnames(carrion_matrix) <- size_bins
# 
# # Convert to a data frame and add species names
# carrion_df <- as.data.frame(carrion_matrix)
# carrion_df$Species <- rownames(carrion_matrix)
# 
# # Reshape the data to long format
# 
# carrion_long <- carrion_df %>%
#   pivot_longer(cols = -Species, names_to = "Size", values_to = "carrion_production") %>%
#   mutate(Size = as.numeric(Size))
# 
# # View the final data frame
# print(carrion_long)



```

## 3b. Approach 2: manually extract species specific external mortality rates and multiply by weight. However, extmort is constant so size is not taken into account.


```{r}
# Function to calculate respired carbon per species and size
carcass_carbon_per_species <- function(params_v1, 
                                        n = params_v1@initial_n, 
                                        n_other = params_v1@initial_n_other) {
    # Get metabolic rates per species and size (g m-2 y-1)
    ext_mort_rates <- getExtMort(params_v1)
    
    # Multiply metabolic rates by abundance for each species and size
    ext_mort_rates <- sweep(ext_mort_rates * n, 2, params_v1@dw, "*", check.margin = FALSE)
    
    # Return the respired carbon per species and size
    return(ext_mort_rates)
}

# Call the function to get respired carbon
ext_mort_rates <- carcass_carbon_per_species(params_v1)

species_names <- rownames(params_v1@species_params)
size_bins <- params_v1@w



# Assign the correct column names (size bins)
colnames(ext_mort_rates) <- size_bins

# Convert to a data frame and add species names
ext_mort_rates_df <- as.data.frame(ext_mort_rates)
ext_mort_rates_df$Species <- rownames(ext_mort_rates)

# Reshape the data to long format

ext_mort_rates_long <- ext_mort_rates_df %>%
  pivot_longer(cols = -Species, names_to = "Size", values_to = "ext_mort_rates") %>%
  mutate(Size = as.numeric(Size))

# View the final data frame
print(ext_mort_rates_long)

# Plot using ggplot2 with log scale for size
CC_p <- ggplot(ext_mort_rates_long, aes(x = Size, y = ext_mort_rates, color = Species)) +
  geom_line() +
  scale_x_log10() +
  labs(title = "Carcass weight (g) per Size for Each Species",
       x = "Size (g)",
       y = "g C m-2 y-1",
       color = "Species") +
  theme_minimal()

```


```{r}


carcass_carbon_df <- ext_mort_rates_long %>% 
  mutate(g_C = ext_mort_rates * 0.4) %>% # change this to species specific carbon content (or leave as assumption - Steinberg et al. 2017?)
  mutate(carcass_C = g_C * 0.1713287/hours_per_year) #constant taken from MizerShelf - proportion of ext_mort that produces carrion (i.e. some of the external mortality will not contribute to carrion production due to consumption from non-modeled species like sea birds)


carrion_p <- ggplot(carcass_carbon_df, aes(x = Size, y = carcass_C, color = as.factor(Species))) +
  geom_line() +
  scale_y_log10()+
  scale_x_log10() +
  labs(title = "Carcass Carbon per Size for Each Species",
       x = "Size (g)",
       y = "g C m-2 h-1",
       color = "Species") +
  theme_minimal()

carrion_p

#ggsave("CC.png", plot = carrion_p)

```



## 3c. Extract external contribution to carrion production (MizerShelf). Ideal approach, but unable to partition natural mortality atm.

```{r}

# #approach 3: extract external contribution to carrion production
# 
# carrion_production_per_species <- function(params_v1, n = params_v1@initial_n,
#                                            rates = getRates(params_v1)) {
#     # Calculate natural mortality contribution to carrion per species
#     mu_b_contribution <- (params_v1@mu_b * n) * (params_v1@w * params_v1@dw) *
#                          params_v1@other_params$carrion$ext_prop
# 
#     # Calculate gear mortality contribution to carrion per species
#     gear_mort_contribution <- (gearMort(params_v1, rates$f_mort) * n) *
#                               (params_v1@w * params_v1@dw)
# 
#     # Calculate discards contribution to carrion per species
#     discards_contribution <- ((rates$f_mort * n) * (params_v1@w * params_v1@dw)) *
#                              params_v1@species_params$discard
# 
#     # Create a list to store carrion production per species
#     carrion_per_species <- list(
#         mu_b = mu_b_contribution,
#         gear_mort = gear_mort_contribution,
#         discards = discards_contribution
#     )
# 
#     return(carrion_per_species)
# }
# 
# # Assume params is already defined and get carrion production per species
# carrion <- carrion_production_per_species(params_v1)
# 
# carrion
# 
# getCarrionProduction(params_v1)


```


# 4. Carbon pools

## 4a. combine carbon products dataframe


```{r}

# First, merge respired_carbon_long and feces_df
combined_df1 <- merge(respired_carbon_df, feces_df, by = c("Species", "Size"))

# Then, merge the result with carcass_carbon_df
combined_df <- merge(combined_df1, carcass_carbon_df, by = c("Species", "Size"))

# View the combined dataframe
head(combined_df)

```

## 4b. add in DVM info to split carbon products up by depth layer - based on vertical_migration_array

```{r}

# Convert vertical_migration_array to a dataframe
vertical_migration_df <- as.data.frame(vertical_migration_array)

# Transpose the dataframe so species names are in a column
vertical_migration_df <- t(vertical_migration_df)

# Convert it back to a dataframe and add row names as a column
vertical_migration_df <- as.data.frame(vertical_migration_df)
vertical_migration_df$Species <- rownames(vertical_migration_df)
rownames(vertical_migration_df) <- NULL


# Rename the columns for easier merging
colnames(vertical_migration_df) <- c("P_Epi", "P_Umeso", "P_Lmeso", "Species")


# Split the Species column into species name and size using strsplit
split_values <- strsplit(as.character(vertical_migration_df$Species), "(?<=\\D)\\.(?=\\d)", perl = TRUE)

# Extract the species name
vertical_migration_df$Species <- sapply(split_values, `[`, 1)

# Extract the size
vertical_migration_df$Size <- sapply(split_values, `[`, 2)
vertical_migration_df <- vertical_migration_df %>% 
  select(-Size)


# Merge the vertical migration data with your combined dataframe
final_df <- merge(combined_df, vertical_migration_df, by = c("Species"), all.x = TRUE) #may need to make this more dynamic and incorporate DVM based on size - to be discussed later

final_df <- final_df %>% 
  unique()
# View the final dataframe
head(final_df)

```



## 4c. Create Carbon pools dataframe

```{r}

DIC_POC <- final_df %>% 
rowwise() %>%
  mutate(total_DIC = sum(respired_carbon),
         total_POC = sum(Fecal_Carbon, carcass_C)) %>%
  ungroup()


carbon_pools <- DIC_POC %>% 
  select(Species, total_DIC, total_POC, P_Epi, P_Umeso, P_Lmeso) %>% 
  group_by(Species) %>% 
  mutate(g_POC_m2_h = sum(total_POC),
         g_DIC_m2_h = sum(total_DIC)) %>% 
  ungroup() %>% 
  select(Species, g_POC_m2_h, g_DIC_m2_h,P_Epi, P_Umeso, P_Lmeso) %>% 
  unique() %>% 
    mutate(DIC_Epi = g_DIC_m2_h *15.8* P_Epi, #based on % of species specific biomass per layer and multiplied by daylight at SOTS during SOLACE (15.8 h)
         DIC_Umeso = g_DIC_m2_h *15.8* P_Umeso,
         DIC_Lmeso = g_DIC_m2_h *15.8* P_Lmeso,
         POC_Epi = g_POC_m2_h *15.8* P_Epi,
         POC_Umeso = g_POC_m2_h *15.8* P_Umeso,
         POC_Lmeso = g_POC_m2_h *15.8* P_Lmeso)



```

###Group by higher taxonomic level

```{r}


# Add the higher_taxa column based on Species
carbon_pools <- carbon_pools %>%
  mutate(higher_taxa = case_when(
    Species == "Atolla wyvillei" ~ "Gelatinous (other)",
    Species == "Gymnosomata" ~ "Gelatinous (other)",
    Species == "Chaetognath" ~ "Gelatinous (other)",
    Species == "Cranchidae" ~ "Squid",
    Species == "Pyrosoma atlanticum" ~ "P. atlanticum",
    Species == "Euphausia similis" ~ "Krill",
    Species == "Thysannoessa macura" ~ "Krill",
    Species == "Bathylagus spp" ~ "Fish",
    Species == "Cyclothone spp" ~ "Fish",
    Species == "Lampanyctus australis" ~ "Fish",
    Species == "Nannobrachium achirus" ~ "Fish",
    Species == "Sergestes spp" ~ "Crustacean (other)",
    Species == "Themisto spp" ~ "Crustacean (other)",
    TRUE ~ NA_character_ # Assign NA for any unmatched species
  ))

# View the updated dataframe
print(carbon_pools)

```


## 4d. Plot respired DIC 

```{r}

# Gather the data into a long format for plotting
DIC_injection_resp_df <- carbon_pools %>%
  select(Species, DIC_Epi, DIC_Umeso, DIC_Lmeso, higher_taxa) %>%
  pivot_longer(cols = c(DIC_Epi, DIC_Umeso, DIC_Lmeso), 
               names_to = "Layer", 
               values_to = "DIC") %>%
  mutate(
    Layer = recode(Layer,
                   "DIC_Epi" = "Epi",
                   "DIC_Umeso" = "Umeso",
                   "DIC_Lmeso" = "Lmeso"),
    Layer = factor(Layer, levels = c("Lmeso", "Umeso", "Epi")),
    total_DIC_mg = DIC * 1000
  )


# Create the horizontal bar plot
DIC_injection_resp_p <- ggplot(DIC_injection_resp_df, aes(x = total_DIC_mg, y = Layer, fill = Species)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "mg C m-2 d-1", y = "Depth Stratum") +
  xlim(0,10)+
  scale_y_discrete(labels = c("Lmeso","Umeso","Epi")) +
  theme(legend.title = element_blank())

DIC_injection_resp_p

ggsave("Supp_fig8_DICpertaxa.png", plot = DIC_injection_resp_p, width = 7, height = 5)
```


```{r}
DIC_injection_total <- DIC_injection_resp_df %>% 
  group_by(Layer) %>% 
  mutate(mg_C_m2_d_layer = sum(total_DIC_mg),
         POC_DIC = "DIC_inject") %>% 
  ungroup() %>% 
  select(Layer, mg_C_m2_d_layer, POC_DIC) %>% 
  unique() 
```



## 4e. Plot POC (fecal & carcass carbon)

```{r}


# Gather the data into a long format for plotting
POC_injection_df <- carbon_pools %>%
  select(Species, POC_Epi, POC_Umeso, POC_Lmeso, higher_taxa) %>%
  pivot_longer(cols = c(POC_Epi, POC_Umeso, POC_Lmeso), 
               names_to = "Layer", 
               values_to = "POC") %>%
  mutate(Layer = factor(Layer, levels = c("POC_Lmeso", "POC_Umeso", "POC_Epi")),
    Layer = recode(Layer,
                   "POC_Epi" = "Epi",
                   "POC_Umeso" = "Umeso",
                   "POC_Lmeso" = "Lmeso"),
    Layer = factor(Layer, levels = c("Lmeso", "Umeso", "Epi")),
    total_POC_mg = POC * 1000
  )

# Create the horizontal bar plot
POC_injection_p <- ggplot(POC_injection_df , aes(x = total_POC_mg, y = Layer, fill = Species)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "mg C m-2 d-1", y = "Depth Stratum") +
  xlim(0,10)+
  scale_y_discrete(labels = c("Lmeso","Umeso","Epi")) +
  theme(legend.title = element_blank())

POC_injection_p

ggsave("Supp_fig9_POCpertaxa.png", plot = POC_injection_p, width = 7, height = 5)


```

#plot higher_taxa carbon pools (Figure 3)

```{r}

#colors (to match Baker et al., 2025)

micronekton_colors <- plasma(6, begin = 0, end = 0.8)   


# Combine micronekton and mesozooplankton colors
taxa_colors <- c(
  "Gelatinous (other)" = micronekton_colors[1],
  "P. atlanticum" = micronekton_colors[2],
  "Krill" = micronekton_colors[3],
  "Squid" = micronekton_colors[4],
  "Crustacean (other)" = micronekton_colors[5],
  "Fish" = micronekton_colors[6]
)

#DIC

DIC_injection_percent_taxa <- DIC_injection_resp_df %>% 
  select(higher_taxa, Layer, total_DIC_mg) %>% 
  group_by(higher_taxa, Layer) %>% 
  mutate(total_DIC_mg = sum(total_DIC_mg)) %>% 
  ungroup() %>% 
  unique() %>% 
  group_by(Layer) %>% 
  mutate(DIC_layer = sum(total_DIC_mg)) %>% 
  ungroup() %>% 
  mutate(percent_cont = total_DIC_mg/DIC_layer)

write.csv(DIC_injection_percent_taxa, "DIC_inject_percent_taxa.csv")

# Create the horizontal bar plot
DIC_injection_highertaxa_p <- ggplot(DIC_injection_resp_df, aes(x = total_DIC_mg, y = Layer, fill = higher_taxa)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "mg C m-2 d-1", y = "Depth Stratum") +
  xlim(0,10)+
  scale_y_discrete(labels = c("Lmeso","Umeso","Epi")) +
  scale_fill_manual(values = taxa_colors, name = NULL)+
  theme(legend.title = element_blank())

DIC_injection_highertaxa_p

```

```{r}
#POC
POC_injection_percent_taxa <- POC_injection_df %>% 
  select(higher_taxa, Layer, total_POC_mg) %>% 
  group_by(higher_taxa, Layer) %>% 
  mutate(total_POC_mg = sum(total_POC_mg)) %>% 
  ungroup() %>% 
  unique() %>% 
  group_by(Layer) %>% 
  mutate(POC_layer = sum(total_POC_mg)) %>% 
  ungroup() %>% 
  mutate(percent_cont = total_POC_mg/POC_layer)

write.csv(POC_injection_percent_taxa, "POC_inject_percent_taxa.csv")

# Create the horizontal bar plot
POC_injection_highertaxa_p <- ggplot(POC_injection_df , aes(x = total_POC_mg, y = Layer, fill = higher_taxa)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "mg C m-2 d-1", y = "Depth Stratum") +
  xlim(0,10)+
  scale_y_discrete(labels = c("Lmeso","Umeso","Epi")) +
  scale_fill_manual(values = taxa_colors, name = NULL)+
  theme(legend.title = element_blank())

POC_injection_highertaxa_p

```


```{r}
# Combine the two plots side by side with a shared legend below
combined_plot <- DIC_injection_highertaxa_p + POC_injection_highertaxa_p +
  plot_layout(guides = "collect", ncol = 2) + # Arrange plots in 2 columns and collect the legend
  plot_annotation(tag_levels = 'A') & # Add tags for panels
  theme(legend.position = "bottom") # Place the legend below

# Display the combined plot
combined_plot

ggsave("Fig3_DIC_POC.png", plot = combined_plot, width = 7, height = 5)

```

```{r}
# Update DIC plot with title and multiline y-axis labels
DIC_injection_highertaxa_p <- DIC_injection_highertaxa_p +
  labs(title = expression(DIC[inject])) +
  scale_y_discrete(labels = c("Epi" = "Epipelagic", 
                              "Umeso" = "Upper\nmesopelagic", 
                              "Lmeso" = "Lower\nmesopelagic")) +
  theme(
    axis.text.y = element_text(angle = 90, hjust = 0.5),
     axis.title.y = element_blank()
  )

# Update POC plot with title, multiline y-axis labels, and remove y-axis title
POC_injection_highertaxa_p <- POC_injection_highertaxa_p +
  labs(title = expression(POC[inject]))+
  scale_y_discrete(labels = c("Epi" = "Epipelagic", 
                              "Umeso" = "Upper\nmesopelagic", 
                              "Lmeso" = "Lower\nmesopelagic")) +
  theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank()
  )

# Combine and save
combined_plot <- DIC_injection_highertaxa_p + POC_injection_highertaxa_p +
  plot_layout(guides = "collect", ncol = 2) &
 # plot_annotation(tag_levels = 'A') & 
  theme(legend.position = "bottom")

combined_plot
ggsave("Fig3_DIC_POC.png", plot = combined_plot, width = 7, height = 5)



```


## Apply microb remin to POC

```{r}

POC_mg_df <- POC_injection_df %>% 
  select(Layer, total_POC_mg) %>% 
  group_by(Layer) %>% 
  mutate(POC_mg = sum(total_POC_mg)) %>% 
  select(Layer, POC_mg) %>% 
  unique()

#find export flux at 200 m, 400 m and 1000 m 
# use F(z) = F(z0)* e^-((K*(z-z0))/w) #Banse 1990
# where F(z) is the flux at depth z, F(z0) is the flux at reference depth, K is remin rate (d-1), w = sinking speed m d-1 and (z-z0) is depth layer

# Define the microbial remineralization rate
#K <- 0.11 # from Louis work
K <- 0.16 #average from Bressac et al. 2024
w <- 8 #Louis work


# POC at upper bounds of each depth layer
POC_Epi <- 4.5   # at 50 m
POC_Umeso <- 3.0 # at 200 m
POC_Lmeso <- 0.66 # at 400 m

# Calculate flux at 200 m (directly from Epi POC)
F_200 <- POC_Epi * exp(-K*(200-50)/w)

# Calculate flux at 400 m using 200 m as reference
F_400 <- (F_200 +POC_Umeso) * exp(-K * (400 - 200)/w)

# Calculate flux at 1000 m using 400 m as reference
F_1000 <- (F_400 + POC_Lmeso) * exp(-K * (1000 - 400)/w)


## calculate DIC remineralised (retained) per layer

# Calculate DIC respired in each layer
DIC_Epi <- POC_Epi - F_200 # we assume all POC sinks below 200 m 
DIC_Umeso <- (POC_Umeso + F_200) - F_400
DIC_Lmeso <- (POC_Lmeso + F_400) - F_1000

# Print results
cat("DIC Respired in Epi layer (0-200 m):", DIC_Epi, "mg\n")
cat("DIC Respired in Umeso layer (200-400 m):", DIC_Umeso, "mg\n")
cat("DIC Respired in Lmeso layer (400-1000 m):", DIC_Lmeso, "mg\n")

# Create a dataframe with the remin DIC values
DIC_remin <- data.frame(
  Layer = c("Epi", "Umeso", "Lmeso"),
  DIC_remin = c(DIC_Epi, DIC_Umeso, DIC_Lmeso), 
  POC_export = c(F_200, F_400, F_1000)
)


```

#combine with DIC injected from respiration

```{r}
DIC_mg_df <- DIC_injection_resp_df %>% 
  select(Layer, total_DIC_mg) %>% 
  group_by(Layer) %>% 
  mutate(DIC_inject = sum(total_DIC_mg)) %>% 
  select(Layer, DIC_inject) %>% 
  unique()

# Add the calculated DIC values to the existing DIC_mg values
DIC_mg_df <- DIC_mg_df %>%
  left_join(DIC_remin, by = "Layer") %>%
  mutate(DIC_mg = DIC_inject + DIC_remin) %>%
  select(Layer, DIC_inject, DIC_mg) # Keep only necessary columns

# Print updated dataframe
print(DIC_mg_df)


```



## How much POC sinks below meso?

```{r}
F_1000

```


#plot carbon pools

```{r}

# Ensure the correct stacking order
DIC_mg_df$Layer <- factor(DIC_mg_df$Layer, levels = c("Epi", "Umeso", "Lmeso"))

# Create a custom blue gradient by replacing the darkest shade
custom_blue_palette <- function(n) {
  pal <- viridis_pal(option = "mako")(n) # Generate the mako palette
  pal[1] <- "#00008B" # Replace the darkest color with a dark blue (hex: DarkBlue)
  return(rev(pal))
}

# Create the stacked bar chart with renamed legend labels
total_DIC_WC <- ggplot(DIC_mg_df, aes(x = "", y = DIC_mg, fill = Layer)) +
  geom_bar(stat = "identity", width = 0.6) +
  scale_fill_manual(
    values = custom_blue_palette(4), # Use the custom blue palette
    labels = c(
      "Epi" = "Epipelagic\n0 - 200 m",
      "Umeso" = "Upper mesopelagic\n200 - 400 m",
      "Lmeso" = "Lower mesopelagic\n400 - 1000 m"
    )
  ) +
  labs(
    title = NULL,
    x = NULL,
    y = "mg C m-² d⁻¹",
    fill = "Layer"
  ) +
  theme_minimal() +
  theme(
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    text = element_text(size = 16),          # Increase overall text size
    axis.title.y = element_text(size = 18),  # Increase y-axis title size
    axis.text.y = element_text(size = 14),   # Increase y-axis text size
    legend.text = element_text(size = 14),   # Increase legend text size
    legend.title = element_text(size = 16)   # Increase legend title size
  )
  

# Save the plot
ggsave("Fig4_DIC_WC.png", plot = total_DIC_WC, width = 7, height = 5)


```


#### make table reflecting new values........


```{r}

total_DIC <- left_join(POC_mg_df, DIC_mg_df, by = "Layer")
total_DIC <- left_join(total_DIC, DIC_remin, by = "Layer")

write.csv(total_DIC, "total_DIC.csv")


```



#Compare values with chapter 1: remove species where % epi = 0 - also need to remove POC sinking contribution.... 

```{r}
#POC_injection_df -= POC with no remineralisation
#DIC_injection_resp_df = respired DIC


# Merge DIC_long and POC_DIC_long by Species and layer
merged_df_no_remin <- left_join(POC_injection_df, DIC_injection_resp_df, by = c("Species", "Layer"))

merged_df_no_remin <- merged_df_no_remin %>%
  mutate(
    Total_DIC = POC + DIC,  # Assuming DIC from DIC_to_POC_df is DIC.x and from POC_DIC_long is DIC.y
  total_DIC_mg = Total_DIC * 1000) %>%   # Convert to milligrams)  
  select(Species, Layer, total_DIC_mg )


merged_df_no_remin <- merged_df_no_remin %>% 
  mutate(mod_type = "mesomizer")



# N. achirus, Cyclothone spp, Bathylagus spp, Gymnosomata, atolla wyvillei

df_no_epi <- merged_df_no_remin  %>% 
  filter(!Species %in% c("Nannobrachium achirus", "Cyclothone spp", "Gymnosomata", "Bathylagus spp", "Atolla wyvillei"))

write.csv(df_no_epi, "modeled_DIC_no_epi.csv")

# Create the horizontal bar plot
Total_DIC_pools_no_remin <- ggplot(df_no_epi, aes(x = total_DIC_mg , y = Layer, fill = Species)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "DIC (mg C m-2 d-1)", y = "Depth Stratum") +
  scale_y_discrete(labels = c("Lmeso","Umeso","Epi")) +
  theme(legend.title = element_blank()) +
  theme(
    axis.text = element_text(size = 14),  # Adjust size of axis text
    axis.title = element_text(size = 16), # Adjust size of axis titles
    legend.text = element_text(size = 12), # Adjust size of legend text
    legend.title = element_text(size = 14) # Adjust size of legend title
  )


# Plot the graph
Total_DIC_pools_no_remin

```





