

```{r}
library(mizerExperimental)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(reshape2)
# remotes::install_github("gustavdelius/mizerShelf")
library(mizerShelf)
#install.packages("remotes")
#remotes::install_github("sizespectrum/therMizer")
library(therMizer)
#install.packages("sensobol")
library(sensobol)
library(sensitivity)
library(dietr)

 
 rm(list =ls()) # clear memory
 
```


```{r}
SIA_TP <- read.csv("POM_meandata.csv")
# Remove all '.' from the Species column in the SIA_TP dataframe
SIA_TP$Species <- gsub("\\.", "", SIA_TP$Species)
# Correct the species name in the trophic_positions_by_size dataframe
SIA_TP$Species <- gsub("Nannobranchium achirus", "Nannobrachium achirus", SIA_TP$Species)


params_v1 <- readRDS("tuned_params_v3.rds")

plotDiet(params_v1)
```



# find trophic position by size
```{r}
# Function to calculate trophic position by size and species
calc_trophic_position_by_size <- function(params, resourceTL = 1.0) {
  # Get diet composition from params object
  dietcomp <- getDiet(params)
  
  # Get species names
  species <- dimnames(dietcomp)$predator
  
   # Set up trophic levels for each prey species (and the basal resource) - from SIA data
  prey_tl <- c(
    "Euphausia similis" = 2.7,  # Example: small zooplankton
    "Thysannoessa macura" = 2.7,
    "Sergestes spp" = 2.8,    # Example: medium zooplankton
    "Themisto spp" = 2.1,
    "Lampanyctus australis" = 3.2,  # Example: small fish
    "Nannobrachium achirus" = 3.2,
    "Cyclothone spp" = 3.4,
    "Bathylagus spp" = 3.3,  # Example: large fish
    "Cranchidae" = 2.7,
    "Chaetognath" = 2.7,       # Example: top predator
    "Gymnosomata" = 3.1,
    "Pyrosoma atlanticum" = 2.0,  # Treat this as basal resource or filter-feeder
    "Atolla wyvillei" = 3.0,  # Example: jellyfish top predator
    "Resource" = resourceTL
  )
  
  
  # Initialize a list to store trophic positions by size for each species
  trophic_positions_by_size <- list()
  
  # Calculate trophic level for each predator species and size class
  for (pred in species) {
    # Get the diet composition for this predator across all size classes and prey
    spec_diet <- dietcomp[pred, , , drop = FALSE]
    
    # Calculate trophic level for each size class
    tl_per_size <- apply(spec_diet, 2, function(diet) {
      weighted_tl <- sum(diet * prey_tl) / sum(diet)  # Weighted mean of prey TL
      return(1 + weighted_tl)  # Add 1 to represent predator being one level above prey
    })
    
    # Store the trophic levels for this species by size class
    trophic_positions_by_size[[pred]] <- tl_per_size
  }
  
  # Convert list to data frame for easier manipulation
  trophic_df <- do.call(rbind, lapply(names(trophic_positions_by_size), function(spec) {
    data.frame(
      Species = spec,
      Size = dimnames(dietcomp)$w,  # Assuming 'w' represents the size classes
      TrophicPosition = trophic_positions_by_size[[spec]]
    )
  }))
  
  return(trophic_df)
}

# Step 2: Apply the function to your params object
trophic_positions_by_size <- calc_trophic_position_by_size(params_v1)

# Step 3: View the trophic positions by size and species
print(trophic_positions_by_size)






```

```{r}

# Manually set the order of the species
species_order <- c(
  "Pyrosoma atlanticum", "Themisto spp", "Euphausia similis", "Thysannoessa macura", "Chaetognath",
  "Cranchidae", "Sergestes spp", "Atolla wyvillei", 
  "Nannobrachium achirus", "Lampanyctus australis", "Bathylagus spp", "Cyclothone spp", "Gymnosomata"
)

# Reorder the Species factor based on the specified order
trophic_positions_by_size$Species <- factor(trophic_positions_by_size$Species, levels = species_order)

# Create the boxplot with specified y-axis limits and ordered species
modeledTP_p <- ggplot(trophic_positions_by_size, aes(x = Species, y = TrophicPosition)) +
  geom_boxplot() + # Adds the boxplot
  labs(x = "Species", y = "Trophic Position", title = "Boxplot of Trophic Position by Species") +
  theme_minimal() + # Apply a minimal theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + # Rotate x-axis labels
  ylim(1, 4) # Set y-axis limits to 0-4


ggsave("modeledTP_p.png", plot = modeledTP_p)

modeledTP_p

```



```{r}

# Step 1: Filter out size classes with near-zero trophic positions
trophic_positions_filtered <- trophic_positions_by_size[trophic_positions_by_size$TrophicPosition > 0, ]

# Step 2: Plot trophic position by size for each species with facet wrap and species-specific x-axis
modeledTP_size <- ggplot(trophic_positions_filtered, aes(x = as.numeric(Size), y = TrophicPosition)) +
  geom_line() +  # Plot lines to show the change in trophic position by size
  theme_minimal() +  # Minimal theme for a clean look
  labs(title = "Trophic Position by Size for Each Species", 
       x = "Size Class", 
       y = "Trophic Position") +
  facet_wrap(~ Species, scales = "free_x") +  # Create facet wrap for each species with free x-axis
  ylim(1, NA) +
  theme(legend.position = "none")  # Remove legend as each facet represents a species

ggsave("modeledTP_size_p.png", plot = modeledTP_size)

modeledTP_size
```
#SIA v. model outputs

```{r}

# Prepare the stable isotope data
SIA_TP_clean <- SIA_TP %>%
  rename(TrophicPosition = TP) %>% # Make column names consistent
  select(Species, TrophicPosition) # Select relevant columns

# Add a column to distinguish the data sources
trophic_positions_by_size$DataSource <- "Modeled"
SIA_TP_clean$DataSource <- "Stable Isotope"


# Combine the two datasets
combined_TP <- bind_rows(trophic_positions_by_size, SIA_TP_clean)

# Manually set the order of the species
species_order <- c(
  "Pyrosoma atlanticum", "Euphausia similis", "Thysannoessa macura",  "Themisto spp", "Sergestes spp", "Chaetognath",
  "Cranchidae","Gymnosomata", "Atolla wyvillei", 
  "Nannobrachium achirus", "Lampanyctus australis", "Bathylagus spp", "Cyclothone spp"
)

# Reorder the Species factor based on the specified order
combined_TP$Species <- factor(combined_TP$Species, levels = species_order)

# Create the combined boxplot
combined_plot <- ggplot(combined_TP, aes(x = Species, y = TrophicPosition, fill = DataSource)) +
  geom_boxplot(position = position_dodge(width = 0.75)) + # Dodge to separate boxes
  labs(x = "Species", y = "Trophic Position", title = "Comparison of Trophic Position by Species") +
  theme_minimal() + # Apply a minimal theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + # Rotate x-axis labels
  ylim(1, 4) + # Set y-axis limits to 1-4
  scale_fill_manual(values = c("Modeled" = "skyblue", "Stable Isotope" = "orange")) # Optional: Set custom colors

# Display the plot
print(combined_plot)

ggsave("modeled_v_SIA.png", plot = combined_plot)

combined_plot

```
#other data vis:

```{r}
ggplot(combined_TP, aes(x = TrophicPosition, fill = DataSource)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ Species) +
  labs(x = "Trophic Position", title = "Density Plot of Trophic Position by Species") +
  theme_minimal()





```


```{r}
# Make sure the Size column is numeric
trophic_positions_by_size$Size <- as.numeric(trophic_positions_by_size$Size)


```
```{r}
trophic_positions_by_size <- trophic_positions_by_size %>% 
  filter(Species != "Pyrosoma atlanticum" & Species != "Euphausia similis" & Species != "Thysannoessa macura") # removed for now 23/10.....

# Fit a linear model
model_w <- lm(TrophicPosition ~ Size, data = trophic_positions_by_size )

# Extract the R-squared value
r_squared <- summary(model_w)$r.squared

# Create the scatterplot with a line of best fit and display the R-squared value
ggplot(trophic_positions_by_size, aes(x = Size, y = TrophicPosition)) +
  geom_point(alpha = 0.6) +  # Scatterplot points with some transparency
  geom_smooth(method = "lm", color = "blue", se = TRUE) +  # Line of best fit with confidence interval
  labs(x = "Size (g)", y = "Trophic Position (TP)", 
       title = "Scatterplot of Trophic Position vs. Size with Line of Best Fit") +
  annotate("text", x = Inf, y = Inf, label = paste("R² =", round(r_squared, 3)), 
           hjust = 1.1, vjust = 1.5, size = 5, color = "black") +  # Add R-squared annotation
  xlim(0, 10) +  # Set the x-axis limit to 0 - 10 g
  theme_minimal()
```


#stats


```{r}
# Two-sample t-test (Welch's t-test)
t_test_results <- t.test(TrophicPosition ~ DataSource, data = combined_TP)
print(t_test_results)

```


```{r}
# Mann-Whitney U test
wilcox_test_results <- wilcox.test(TrophicPosition ~ DataSource, data = combined_TP)
print(wilcox_test_results)

```

```{r}
# Two-way ANOVA
anova_results <- aov(TrophicPosition ~ Species * DataSource, data = combined_TP)
summary(anova_results)

```

```{r}
# Linear mixed-effects model
library(lme4)
model <- lmer(TrophicPosition ~ DataSource + (1 | Species), data = combined_TP)
summary(model)

```
The model shows that there is a significant difference in trophic position between the modeled and stable isotope-derived data, with stable isotope values generally being higher. The random effects account for species-specific variability, showing that while species influence trophic positions, the difference between data sources remains a strong effect.


