---
title: "03_MesoMizer_shelf"
output: html_document
date: "2024-06-05"
---



```{r}
library(mizerExperimental)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
 remotes::install_github("gustavdelius/mizerShelf")
library(mizerShelf)
 
 rm(list =ls()) # clear memory
 
```

#Parameter description--------------------------------------------------

```{r}

param_description <- data.frame("parameter" = c("w_inf","w_mat","beta","sigma","R_max","k_vb","l25","l50",
                                                "a","b","catchability","h","k","ks","z0","gamma","w_mat25","erepro"),
                                "description" = c("asymptotic weigth",
                                                  "maturation weight (determines when 50% of the population has matured using a sigmoid function)",
                                                  "preferred predator/prey mass ratio","width of the feeding kernel",
                                                  "Beverton-Holt density dependence parameter",
                                                  "von Bertalanffy growth parameter","length at 25% catch","length at 50% catch",
                                                  "coefficient for age to size conversion",
                                                  "constant for age to size conversion","fisheries efficiency","maximum intake rate","metabolism constant",
                                                  "metabolism coefficient","background mortality coefficient",
                                                  "search volume (obtained from beta and sigma)",
                                                  "weight at which 25% of individuals are mature","reproductive output scalar")
) #add biomass observed 

knitr::kable(param_description, caption = "parameters' description")


```

#Species parameters


```{r}
sp_1 <-
  data.frame(
    "species" = c("Euphausia similis","Thysannoessa macura","Sergestes spp.", "Themisto spp", "Lampanyctus australis","Nannobrachium achirus","Cyclothone spp.","Bathylagus spp.","Cranchidae","Chaetognath","Gymnosomata","Pyrosoma atlanticum","Atolla wyvillei"),
    "w_inf" = c(0.417,0.318,1.7,0.1,26.78,58.3,1.62,81.43,20,3.6,0.28,375.8,12),
    "beta" = c(log10(7.2),log10(7.2),log10(2),log10(2),log10(2),log10(2),log10(2),log10(2),log10(1.4),log10(1.35),log10(1.35),log10(7.7),log10(3.7)),
    "biomass_observed" = c( 1.714941e+11,1.561756e+11,3.815464e+10,6.383045e+10,4.760506e+11,2.213141e+10,1.499316e+11,5.021544e+10,1.675264e+11,3.649106e+12,4.251151e+09,3.002268e+13,7.149292e+10),
    "k_vb" = c(0.6, 0.4, 0.8, 0.6, 0.2, 0.4, 0.9, 0.5, 0.4, 0.2, 0.2,0.2,0.2),
    "gear_mort" = c(0,0,0,0,0,0,0,0,0,0,0,0,0),
    "RQ" = c(0.9, 0.9, 1.0, 0.9, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 0.8, 1.0),
    "p_FC" = c(0.4, 0.4, 0.4,0.4, 0.4, 0.4,0.4, 0.4, 0.4,0.4, 0.4, 0.4,0.4), # carbon content of fecal pellets (%)
    "z0" = c(0.1, 0.1, 0.1,0.1, 0.1, 0.1, 0.1, 0.1, 0.1,0.1, 0.1, 0.1, 0.1), # carrion parameter - may delete
    "rho" = c(200, 200, 200, 200, 200, 200,200, 200, 200,200, 200, 200,200) # carrion parameter - may delete
  )
#"sigma" = c(0.7,0.7, 1.3,1.3,1.3,1.3,1.3,1.3,1.3,0.46,0.57,0.7,0.52)) - this causes p. atlanticum to go extinct(too narrow... factor these in later)

              
knitr::kable(sp_1, caption = "species' parameters")


                
                   
```

#Species interaction matrix

```{r}

sp_1_inter <- matrix (c(1,	1,	0.5,	0,	0,	0,	0.5,	0.5,	1,	0,	1,	0,	1,
1,	1,	0.5,	0,	0,	0,	0.5,	0.5,	1,	0,	1,	0,	1,
0.5,	0.5,	1,	0.5,	0.5,	0.5,	1,	1,	0.5,	0.5,	0.5,	0.5,	0.5,
0,	0,	0.5,	1,	0.5,	0.5,	0.08,	0.08,	0,	1,	0,	1,	0,
0,	0,	0,	0.5,	1,	1,	0.58,	0.58,	0,	0.5,	0,	0.5,	0,
0,	0,	0,	0.5,	1,	1,	0.58,	0.58,	0,	0.5,	0,	0.5,	0,
0.5,	0.5,	0.5,	0.08,	0.58,	0.58,	1,	1,	0.5,	0.08,	0.5,	0.08,	0.5,
0.5,	0.5,	0.5,	0.08,	0.58,	0.58,	1,	1,	0.5,	0.08,	0.5,	0.08,	0.5,
1,	1,	0.5,	0,	0,	0,	0.5,	0.5,	1,	0,	1,	0,	1,
0,	0,	0.5,	1,	0.5,	0.5,	0.08,	0.08,	0,	1,	0,	1,	0,
1,	1,	0.5,	0,	0,	0,	0.5,	0.5,	1,	0,	1,	0,	1,
0,	0,	0.5,	1,	0.5,	0.5,	0.08,	0.08,	0,	1,	0,	1,	0,
1,	1,	0.5,	0,	0,	0,	0.5,	0.5,	1,	0,	1,	0,	1

), ncol = 13,
                     dimnames = list(sp_1$species, sp_1$species) )

knitr::kable(sp_1_inter, caption = "species' interactions")

```

```{r}
sp_1<- newMultispeciesParams(sp_1, sp_1_inter, w_pp_cutoff = 1 ) #not sure what w_pp_cutoff - look this up....
species_params(sp_1)
```
```{r}

meso_p1 <- steady(sp_1)

params_v1 <- tuneParams(sp_1)
```

# Respiration

```{r}
#Get metabolic rate per species and size (g y-1)
metab(params_v1)


# Assume params is already defined and get metabolic rates per species and size
metabolic_rates <- metab(params_v1)

# Convert metabolic rates array to long-format data frame
metabolic_rates_df <- melt(metabolic_rates, varnames = c("Species", "Size"), value.name = "Metabolic_Rate")

# Plot using ggplot2 with log scale for size
ggplot(metabolic_rates_df, aes(x = Size, y = Metabolic_Rate, color = as.factor(Species))) +
  geom_line() +
  scale_x_log10() +
  labs(title = "Metabolic Rate per Size for Each Species",
       x = "Size (w)",
       y = "Metabolic Rate",
       color = "Species") +
  theme_minimal()
```
# Convert respiration rate to respired carbon (mg C)


```{r}
#equation: RC = metabolic_rates_df$Metabolic_Rate * RQ * (12/22.4)
#note - numbers will be off. Need to find species specific oxygen intake values for k parameter. Right now metabolic rate does not reflect oxygen consumption so conversion to C is not correct.

# Define the function to convert metabolic rates to respired carbon
metab_RC <- function(metabolic_rates_df, params) {
  # Extract the respiratory quotient (RQ) from the species parameters in params
  RQ <- params@species_params$RQ
  
  # Calculate respired carbon (RC) for each species and size
  metabolic_rates_df$Respired_Carbon <- metabolic_rates_df$Metabolic_Rate * RQ * (12 / 22.4)
  
  return(metabolic_rates_df)
}


species_names <- rownames(params_v1@species_params)
size_bins <- params_v1@w

# Create a data frame with species names, sizes, and metabolic rates
metabolic_rates_df <- expand.grid(Species = species_names, Size = size_bins)
metabolic_rates_df$Metabolic_Rate <- as.vector(metab(params_v1))

# Convert metabolic rates to respired carbon
respired_carbon_df <- metab_RC(metabolic_rates_df, params_v1)

# View the resulting data frame
head(respired_carbon_df)


```
## plot

```{r}
# Plot using ggplot2 with log scale for size
RC_p <- ggplot(respired_carbon_df, aes(x = Size, y = Respired_Carbon, color = as.factor(Species))) +
  geom_line() +
  scale_x_log10() +
  labs(title = "Respired Carbon per Size for Each Species",
       x = "Size (w)",
       y = "RC (g C y-1)",
       color = "Species") +
  theme_minimal()

ggsave(filename = "RC_per_size_species.jpeg", plot = RC_p, width = 8, height = 6, dpi = 300)
```

#simulation - Respired carbon over time (steady state)


```{r}
sim_v1 <- project(params_v1, t_max = 15, t_save = 0.1)

# Function to extract metabolic rates and convert to respired carbon over time

extract_and_convert_rc <- function(sim, params) {
  time_steps <- dim(sim@n)[1]
  species_names <- rownames(params@species_params)
  size_bins <- params@w
  
  # Initialize an empty list to store RC data frames for each time step
  RC_list <- vector("list", time_steps)
  
  for (i in 1:time_steps) {
    n <- sim@n[i, , ]
    n_other <- sim@n_other[i, ]
    
    # Extract rates using the current abundance
    rates <- getRates(params, n = n, n_other = n_other)
    metabolic_rates <- metab(params)
    
    # Create a data frame with species names, sizes, and metabolic rates
    metabolic_rates_df <- expand.grid(Species = species_names, Size = size_bins)
    metabolic_rates_df$Metabolic_Rate <- as.vector(metabolic_rates)
    
    # Convert metabolic rates to respired carbon
    respired_carbon_df <- metab_RC(metabolic_rates_df, params)
    respired_carbon_df$Time <- (i - 1) * 0.1  # Adjust according to t_save
    
    RC_list[[i]] <- respired_carbon_df
  }
  
  # Combine all the data frames in the list into one data frame
  RC_over_time_df <- do.call(rbind, RC_list)
  
  return(RC_over_time_df)
}

# Extract and convert RC over time
RC_over_time_df <- extract_and_convert_rc(sim_v1, params_v1)

```
# sum RC for all size classes per species

```{r}

RC_over_time_df <- RC_over_time_df %>% 
  group_by(Species, Time) %>% 
  mutate(RC_g_C_y = sum(Respired_Carbon)) %>% 
  ungroup() %>% 
  select(Species, RC_g_C_y, Time) %>% 
  unique()

```


#plot RC over time per species

```{r}
RC_over_time_p <- ggplot(RC_over_time_df, aes(x = Time, y = RC_g_C_y, color = Species, group = Species)) +
  geom_line() +
  labs(title = "Respired Carbon (RC) per Species",
       x = "Time",
       y = "g C y-1",
       color = "Species") +
  theme_minimal()

ggsave(filename = "RC_over_time_p.jpeg", plot = RC_over_time_p, width = 8, height = 6, dpi = 300)


```


# Fecal Carbon

```{r}
detritus_production_per_species <- function(params_v1, n = params_v1@initial_n, # in MizerShelf, detritus production includes feces, decomposing carrion and external. For now we only want to extract feces, but we'll keep the other functions in at this stage.
                                            n_other = params_v1@initial_n_other,
                                            rates = getRates(params_v1)) {
    # Calculate the consumption for each species
    consumption <- sweep((1 - rates$feeding_level) * rates$encounter * n, 2,
                         params_v1@dw, "*", check.margin = FALSE)
    
    # Calculate the feces production for each species
    feces <- sweep(consumption, 1, (1 - params_v1@species_params$alpha), "*", 
                   check.margin = FALSE)
    
    # Calculate the carrion decomposition
    carrion <- params_v1@other_params$carrion$decompose * n_other$carrion
    
    # Create a list to store detritus production per species
    detritus_per_species <- list(
        external = params_v1@other_params$detritus$external,
        feces = feces,
        carrion = carrion
    )
    
    return(detritus_per_species)
}

detritus_production_per_species(params_v1)

```



#plot

```{r}

detritus <- detritus_production_per_species(params_v1)

# Convert feces matrix to a long-format data frame
feces_df <- melt(detritus$feces, varnames = c("Species", "Size"), value.name = "Detritus")


FC_p <- ggplot(feces_df, aes(x = Size, y = Detritus, color = as.factor(Species))) +
  geom_line() +
  scale_y_log10()+
  scale_x_log10() +
  labs(title = "Fecal pellet Production per Size for Each Species",
       x = "Size (w)",
       y = "Detritus Production",
       color = "Species") +
  theme_minimal()


ggsave(filename = "FC_per_size_species.jpeg", plot = RC_p, width = 8, height = 6, dpi = 300)

```
# Convert feces (g y-1) to (g C y-1)

```{r}
#check feces is g y-1
#FC = feces_df$Detritus * P_fc

# Extract the P_fc parameter from params_v1
p_FC <- params_v1@species_params$p_FC


# Ensure P_fc is a named vector with species names as names
names(p_FC) <- rownames(params_v1@species_params)

# Create a data frame from P_fc
P_fc_df <- data.frame(Species = names(p_FC), p_FC = p_FC)

# Join the p_FC values with the feces_df dataframe based on Species
feces_df <- feces_df %>%
  left_join(P_fc_df, by = "Species")

# Calculate Fecal Carbon (FC)
feces_df <- feces_df %>%
  mutate(Fecal_Carbon = Detritus *p_FC)

# View the resulting feces_df
head(feces_df)

ggplot(feces_df, aes(x = Size, y = Fecal_Carbon, color = as.factor(Species))) +
  geom_line() +
  scale_y_log10()+
  scale_x_log10() +
  labs(title = "Fecal Carbon per Size for Each Species",
       x = "Size (w)",
       y = "g C y-1",
       color = "Species") +
  theme_minimal()

```

# Simulation of fecal carbon production over time (steady state)

```{r}

# Define the function to convert detritus to fecal carbon
convert_to_fecal_carbon <- function(detritus_df, params) {
  # Extract the P_fc parameter from the species parameters in params
  p_FC <- params@species_params$p_FC
  
  # Ensure P_fc is a named vector with species names as names
  names(p_FC) <- rownames(params@species_params)
  
  # Merge the P_fc values with the detritus_df dataframe based on Species
  detritus_df <- detritus_df %>%
    left_join(data.frame(Species = names(p_FC), p_FC = p_FC), by = "Species")
  
  # Calculate Fecal Carbon (FC)
  detritus_df <- detritus_df %>%
    mutate(Fecal_Carbon = Detritus * p_FC)
  
  return(detritus_df)
}

# Define the function to extract detritus and convert it to fecal carbon over time
extract_and_convert_fc <- function(sim, params) {
  time_steps <- dim(sim@n)[1]
  species_names <- rownames(params@species_params)
  size_bins <- params@w
  
  # Initialize an empty list to store FC data frames for each time step
  FC_list <- vector("list", time_steps)
  
  for (i in 1:time_steps) {
    n <- sim@n[i, , ]
    n_other <- sim@n_other[i, ]
    
    # Extract rates using the current abundance
    rates <- getRates(params, n = n, n_other = n_other)
    
    # Calculate detritus production for each species and size
    feces <- sweep((1 - rates$feeding_level) * rates$encounter * n, 2,
                   params@dw, "*", check.margin = FALSE)
    detritus_df <- expand.grid(Species = species_names, Size = size_bins)
    detritus_df$Detritus <- as.vector(feces)
    
    # Convert detritus to fecal carbon
    fecal_carbon_df <- convert_to_fecal_carbon(detritus_df, params)
    fecal_carbon_df$Time <- (i - 1) * 0.1  # Adjust according to t_save
    
    FC_list[[i]] <- fecal_carbon_df
  }
  
  # Combine all the data frames in the list into one data frame
  FC_over_time_df <- do.call(rbind, FC_list)
  
  return(FC_over_time_df)
}



# Example parameters object (params_v1) initialization
# Make sure your params_v1 is correctly initialized as a MizerParams object

# Run the simulation
sim_v1 <- project(params_v1, t_max = 15, t_save = 0.1)

# Extract and convert FC over time
FC_over_time_df <- extract_and_convert_fc(sim_v1, params_v1)



```


# sum FC for all size classes per species

```{r}

FC_over_time_df <- FC_over_time_df %>% 
  group_by(Species, Time) %>% 
  mutate(FC_g_C_y = sum(Fecal_Carbon)) %>% 
  ungroup() %>% 
  select(Species, FC_g_C_y, Time) %>% 
  unique()

```


#plot RC over time per species

```{r}
FC_over_time_p <- ggplot(FC_over_time_df, aes(x = Time, y = FC_g_C_y, color = Species, group = Species)) +
  geom_line() +
  labs(title = "Fecal Carbon (FC) per Species",
       x = "Time",
       y = "g C y-1",
       color = "Species") +
  theme_minimal()

ggsave(filename = "FC_over_time_p.jpeg", plot = FC_over_time_p, width = 8, height = 6, dpi = 300)

```



#Carrion production

```{r}
carrion_production_per_species <- function(params_v1, n = params_v1@initial_n, 
                                           rates = getRates(params_v1)) {
    # Calculate natural mortality contribution to carrion per species
    mu_b_contribution <- (params_v1@mu_b * n) * (params_v1@w * params_v1@dw) *
                         params_v1@other_params$carrion$ext_prop
    
    # Calculate gear mortality contribution to carrion per species
    gear_mort_contribution <- (gearMort(params_v1, rates$f_mort) * n) * 
                              (params_v1@w * params_v1@dw)
    
    # Calculate discards contribution to carrion per species
    discards_contribution <- ((rates$f_mort * n) * (params_v1@w * params_v1@dw)) *
                             params_v1@species_params$discard
    
    # Create a list to store carrion production per species
    carrion_per_species <- list(
        mu_b = mu_b_contribution,
        gear_mort = gear_mort_contribution,
        discards = discards_contribution
    )
    
    return(carrion_per_species)
}

# Assume params is already defined and get carrion production per species
carrion <- carrion_production_per_species(params_v1)

carrion

getCarrionProduction(params_v1)
getCarrionConsumption(params_v1)

```




#plot - need to work out how to get natural mortality

```{r}
# # Convert mu_b, gear_mort, and discards matrices to long-format data frames
# mu_b_df <- melt(carrion$mu_b, varnames = c("Species", "Size"), value.name = "Carrion")
# gear_mort_df <- melt(carrion$gear_mort, varnames = c("Species", "Size"), value.name = "Carrion")
# discards_df <- melt(carrion$discards, varnames = c("Species", "Size"), value.name = "Carrion")
# 
# # Combine data frames and add a Type column
# mu_b_df$Type <- "mu_b"
# gear_mort_df$Type <- "gear_mort"
# discards_df$Type <- "discards"
# 
# carrion_df <- rbind(mu_b_df, gear_mort_df, discards_df)
# 
# # Plot using ggplot2 with log scale for size
# ggplot(carrion_df, aes(x = Size, y = Carrion, color = as.factor(Species))) +
#   geom_line() +
#   facet_wrap(~ Type) +
#   scale_x_log10() +
#   labs(title = "Carrion Production per Size for Each Species",
#        x = "Size (w)",
#        y = "Carrion Production",
#        color = "Species") +
#   theme_minimal()
```





--------------------------------------------------------------------------------------------


#Carbon Injection potential - need to integrate carrion at later stage.


```{r}
# Function to calculate CIP production per species and size
CIP_production <- function(params_v1, n = params_v1@initial_n,
                           n_other = params_v1@initial_n_other,
                           rates = getRates(params_v1)) {
    # Get detritus production
    detritus <- detritus_production_per_species(params_v1, n, n_other, rates)
    
    # Calculate total detritus production (sum feces and carrion)
    total_detritus <- detritus$feces
    
    # Get metabolic rates
    metabolic_rates <- metab(params_v1)
    
    # Calculate CIP production
    CIP <- total_detritus + metabolic_rates # need to change this to RC
    
    return(CIP)
}

# Example usage
# Assume params is already defined
CIP <- CIP_production(params_v1)

# Convert CIP array to long-format data frame for plotting
CIP_df <- melt(CIP, varnames = c("Species", "Size"), value.name = "CIP_Production")

# Plot using ggplot2 with log scale for size

ggplot(CIP_df, aes(x = Size, y = CIP_Production, color = as.factor(Species))) +
  geom_line() +
  scale_y_log10()+
  scale_x_log10() +
  labs(title = "CIP Production per Size for Each Species",
       x = "Size (w)",
       y = "g C y-1",
       color = "Species") +
  theme_minimal()
```









