---
title: "05_MesoMizer_Shelf_ther"
output: html_document
date: "2024-08-13"
---

This script contains MizerShelf and therMizer packages.


```{r}
library(mizerExperimental)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(reshape2)
# remotes::install_github("gustavdelius/mizerShelf")
library(mizerShelf)
#install.packages("remotes")
#remotes::install_github("sizespectrum/therMizer")
library(therMizer)
#install.packages("sensobol")
library(sensobol)
library(sensitivity)
library(dietr)
library(rgl)
library(viridis)
library(patchwork)
library(scales)

 
 rm(list =ls()) # clear memory
 
```

#Parameter description--------------------------------------------------

```{r}

param_description <- data.frame("parameter" = c("w_inf","w_mat","beta","sigma","R_max","k_vb","l25","l50",
                                                "a","b","catchability","h","k","ks","z0","gamma","w_mat25","erepro", "RQ", "gear_mort", "p_FC", "z0", "rho"),
                                "description" = c("asymptotic weigth",
                                                  "maturation weight (determines when 50% of the population has matured using a sigmoid function)",
                                                  "preferred predator/prey mass ratio","width of the feeding kernel",
                                                  "Beverton-Holt density dependence parameter",
                                                  "von Bertalanffy growth parameter","length at 25% catch","length at 50% catch",
                                                  "coefficient for age to size conversion",
                                                  "constant for age to size conversion","fisheries efficiency","maximum intake rate","metabolism constant",
                                                  "metabolism coefficient","background mortality coefficient",
                                                  "search volume (obtained from beta and sigma)",
                                                  "weight at which 25% of individuals are mature","reproductive output scalar", "respiratory quotient", "mortality rate due to fishing gear", "carbon content of fecal pellets (%)", "carrion parameter", "carrion parameter")
) #add biomass observed 

knitr::kable(param_description, caption = "parameters' description")


```

#Species parameters


```{r}
sp_1 <-
  data.frame(
    "species" = c("Euphausia similis","Thysannoessa macura","Sergestes spp", "Themisto spp", "Lampanyctus australis","Nannobrachium achirus","Cyclothone spp","Bathylagus spp","Cranchidae","Chaetognath","Gymnosomata","Pyrosoma atlanticum","Atolla wyvillei"),
    "w_inf" = c(0.417,0.318,1.7,0.1,26.78,58.3,3.0,81.43,20,1.0,0.28,375.8,12), #check cyclothone in fishbase & chaetognath in sealifebase 
    "beta" = c(15849000, 15849000, 100, 100, 100, 100, 100, 100, 25, 22, 22, 50119000, 5012), #lower ppmr for piscivores, higher for omnivores and planktivores
    "biomass_observed" = c(0.1970, 0.1582, 0.0704,0.1012,0.6665,0.0658,0.4307,0.1493,0.2333,4.3799,0.0126,31.4392,0.1215), #biomass in g m-2. See Biomass_SOLACE.rmd
    "k_vb" = c(0.6, 0.4, 0.8, 0.6, 0.2, 0.4, 0.9, 0.5, 0.4, 0.2, 0.2,0.2,0.2),
    "gear_mort" = c(0,0,0,0,0,0,0,0,0,0,0,0,0),
    "RQ" = c(0.9, 0.9, 1.0, 0.9, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 0.8, 1.0), #0.8 - herbivores, 0.9 omnivores and 1 carnivores
    "p_FC" = c(0.4, 0.4, 0.4,0.4, 0.4, 0.4,0.4, 0.4, 0.4,0.4, 0.4, 0.4,0.4), # carbon content of fecal pellets (%)
    "z0" = c(0.1, 0.1, 0.1,0.1, 0.1, 0.1, 0.1, 0.1, 0.1,0.1, 0.1, 0.1, 0.1), # carrion parameter - may delete
   "rho" = c(200, 200, 200, 200, 200, 200,200, 200, 200,200, 200, 200,200)) # carrion parameter - may delete
#"sigma" = c(0.7,0.7, 1.3,1.3,1.3,1.3,1.3,1.3,1.3,0.46,0.57,0.7,0.52)) #- this causes p. atlanticum to go extinct(too narrow... factor these in later)

              
knitr::kable(sp_1, caption = "species' parameters")


                
                   
```


# Species interaction matrix

### add in DVM based on % of community at each depth
```{r}
#predator = row, prey = column
#sp_2_inter reflects feeding type and DVM
sp_2_inter <-matrix <- matrix(c(
  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
  0.6700000000000002, 0.55, 1, 0.8900000000000001, 0, 0, 0, 0, 0, 0.66, 0.35, 0.65, 0.4,
  0.78, 0.6599999999999999, 0.8900000000000001, 1, 0, 0, 0, 0, 0, 0.7699999999999999, 0.29, 0.6000000000000001, 0.35,
  0.42, 0.30000000000000004, 0.62, 0.57, 0.9999999999999999, 0.22, 0.304, 0.22, 0.45, 0.41000000000000003, 0.22, 0.78, 0.7799999999999999,
  0.07, 0.03, 0.35, 0.29, 0.22, 1, 0.916, 1, 0.2, 0.13, 1, 0.03, 0.21,
  0.15400000000000003, 0.114, 0.434, 0.374, 0.304, 0.916, 1, 0.916, 0.28400000000000003, 0.21400000000000002, 0.916, 0.114, 0.238,
  0.07, 0.03, 0.35, 0.29, 0.22, 1, 0.916, 1, 0.2, 0.13, 1, 0.03, 0.21,
  0.8700000000000001, 0.8300000000000001, 0.7, 0.81, 0.45, 0.2, 0.28400000000000003, 0.2, 1, 0.93, 0.2, 0.5, 0.24000000000000002,
  0.9300000000000002, 0.8900000000000001, 0.66, 0.7699999999999999, 0.41000000000000003, 0.13, 0.21400000000000002, 0.13, 0.93, 1, 0.13, 0.53, 0.2,
  0.07, 0.03, 0.35, 0.29, 0.22, 1, 0.916, 1, 0.2, 0.13, 1, 0.03, 0.21,
  0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
  0.21000000000000002, 0.09, 0.4, 0.35, 0.7799999999999999, 0.21, 0.238, 0.21, 0.24000000000000002, 0.2, 0.21, 0.5700000000000001, 1
), nrow = 13, byrow = TRUE,
  dimnames = list(sp_1$species, sp_1$species) )

knitr::kable(sp_2_inter, caption = "species' interactions")

# Convert the interaction matrix to a data frame
sp_2_inter_df <- as.data.frame(sp_2_inter)

# Ensure row names are preserved as a column (optional)
sp_2_inter_df$Predator <- rownames(sp_2_inter_df)

# Rearrange columns so 'Predator' is the first column (optional)
#sp_2_inter_df <- sp_2_inter_df[, c(ncol(sp_2_inter_df), 1:(ncol(sp_2_inter_df) - 1))]

write.csv(sp_2_inter_df, "sp_2_interactions.csv", row.names = FALSE)

# Print the data frame using knitr::kable for display
knitr::kable(sp_2_inter_df, caption = "Species' interactions")


```



# Tune params

```{r}
 
 # sp_2 <- newMultispeciesParams(sp_1, sp_2_inter, w_pp_cutoff = 0.01)
 # params_v1 <- setParams(sp_2)
 # params_v1 <- tuneParams(sp_2)

params_v0 <- readRDS("tuned_params_DVM3.rds")
params_v0 <- tuneParams(params_v0)


```




# Thermizer -> set thermal tolerances for species

```{r}

# Assign them thermal tolerance limits - general thresholds - will look these up
temp_min <- c(-5, -5,-5, -5,-5, -5,-5, -5,-5, -5,-5, 0,-5)
temp_max <- c(13, 13,13, 13,13, 13,13, 13,13, 12,13, 20,13) #revisit
species_params(params_v0)$temp_min <- temp_min
species_params(params_v0)$temp_max <- temp_max
```


### set temperature parameters for DVM

```{r}
realm_names <- c("Epi","Umeso","Lmeso")
species_names <- as.character(params_v0@species_params$species)
sizes <- params_v0@w

# Create the vertical migration array and fill it
vertical_migration_array <- array(0, dim = (c(length(realm_names), 
                                  length(species_names), length(sizes))), 
                                  dimnames = list(realm = realm_names, sp = species_names, 
                                  w = signif(sizes,3))) # realm x species x size

Epi <- which(realm_names == "Epi")
Umeso <- which(realm_names == "Umeso") 
Lmeso <- which(realm_names == "Lmeso") 

# Define a list of species and their corresponding time allocations - based on average % per depth from SOLACE (Biomass_SOLACE.rmd)
species_allocations <- list(
  "Lampanyctus australis" = c(Epi = 0.21, Umeso = 0.57, Lmeso = 0.22),
  "Euphausia similis" = c(Epi = 0.79, Umeso = 0.14, Lmeso = 0.07),
  "Thysannoessa macura" = c(Epi = 0.91, Umeso = 0.06, Lmeso = 0.03),
  "Sergestes spp" = c(Epi = 0.46, Umeso = 0.19, Lmeso = 0.35),
  "Themisto spp" = c(Epi = 0.57, Umeso = 0.14, Lmeso = 0.29),
  "Nannobrachium achirus" = c(Epi = 0, Umeso = 0, Lmeso = 1),
  "Cyclothone spp" = c(Epi = 0.056, Umeso = 0.028, Lmeso = 0.916), 
  "Bathylagus spp" = c(Epi = 0, Umeso = 0, Lmeso = 1),
  "Cranchidae" = c(Epi = 0.76, Umeso = 0.04, Lmeso = 0.20),
  "Chaetognath" = c(Epi = 0.80, Umeso = 0.07, Lmeso = 0.13),
  "Gymnosomata" = c(Epi = 0, Umeso = 0, Lmeso = 1),
  "Pyrosoma atlanticum" = c(Epi = 0.43, Umeso = 0.54, Lmeso = 0.03),
  "Atolla wyvillei" = c(Epi = 0, Umeso = 0.79, Lmeso = 0.21)
)

# Loop through the species and populate the vertical migration array
for (sp in names(species_allocations)) {
  sp_index <- which(species_names == sp)
  allocations <- species_allocations[[sp]]
  vertical_migration_array[Epi, sp_index, ] <- allocations["Epi"]
  vertical_migration_array[Umeso, sp_index, ] <- allocations["Umeso"]
  vertical_migration_array[Lmeso, sp_index, ] <- allocations["Lmeso"]
}

```


### set up exposure array

```{r}
exposure_array <- array(0, dim = (c(length(realm_names), length(species_names))), 
                  dimnames = list(realm = realm_names, sp = species_names)) # realm x species

for (r in seq(1,length(realm_names),1)) {
    for (s in seq(1,length(species_names),1)) {
        if (any(vertical_migration_array[r,s,] > 0)) {
            exposure_array[r,s] = 1
        }
    }
}


# Create temperature array with constant values per depth layer
times <- 0:500
ocean_temp_array <- array(NA, dim = c(length(times), length(realm_names)), 
                          dimnames = list(time = times, realm = realm_names))

# Set constant temperatures for each realm
for (i in 1:length(times)) {
  ocean_temp_array[i,] <- c(10.5, 10, 7.5) # 10.5°C for Epi, 10°C for Umeso, 7.5°C for Lmeso
}

```


### upgrade therParams - (incorporate temp)

```{r}


params_v1 <- upgradeTherParams(params = params_v0, 
                            temp_min = temp_min,
                            temp_max = temp_max,
                            ocean_temp_array = ocean_temp_array,
                           # n_pp_array = n_pp_array, 
                            vertical_migration_array = vertical_migration_array,
                            exposure_array = exposure_array, 
                            aerobic_effect = TRUE, 
                            metabolism_effect = TRUE)
                                

species_params_df <- params_v1@species_params
write.csv(species_params_df, file = "species_params.csv", row.names = FALSE)


```


#Supplementary plots---------------------------------------------------------------------

```{r}

sim_2 <- project(params_v1)




plotSpectra(params_v1, power = 0, total = TRUE) + 
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),  # Adjust size of axis text
    axis.title = element_text(size = 16), # Adjust size of axis titles
    legend.text = element_text(size = 12), # Adjust size of legend text
    legend.title = element_text(size = 14) # Adjust size of legend title
  )


ggsave("Supp_fig1_ss0.png")


plotSpectra(params_v1, power = 1, total = TRUE) + 
  theme_minimal() +
  theme(
    axis.text = element_text(size = 14),  # Adjust size of axis text
    axis.title = element_text(size = 16), # Adjust size of axis titles
    legend.text = element_text(size = 12), # Adjust size of legend text
    legend.title = element_text(size = 14) # Adjust size of legend title
  )


ggsave("Supp_fig2_ss1.png")


plotSpectra(params_v1, power = 2, total = TRUE) + theme_minimal() #slope should be close to zero - biomass constant over log size bins

ggsave("Supp_fig3_ss2.jpeg")

plotDiet(params_v1)
ggsave("Supp_fig4_diet.png")

plotDeath(params_v1, proportion = T)   + theme_minimal()
ggsave("Supp_fig5_death.png")


plotBiomassObservedVsModel(params_v1, labels = FALSE) + theme_minimal()
ggsave("Supp_fig6_biomassobservedvmodel.png")


plotGrowthCurves(params_v1, species_panel = T)  + theme_minimal()
ggsave("Supp_fig7_growth_curves.png")

plotEnergyBudget(params_v1)

slope_biomass <- getCommunitySlope(sim_2)

slope_biomass[1,]
slope_biomass[100,]

plot(slope_biomass)

```



# Carbon outputs

### 1. Respiratory carbon (DIC)

#### 1a. Get metabolic rates per species & size

```{r}
#Get metabolic rate per species and size (g m-2 y-1)
metabolic_rates <- metab(params_v1)

# Convert metabolic rates array to long-format data frame
metabolic_rates_df <- melt(metabolic_rates, varnames = c("Species", "Size"), value.name = "Metabolic_Rate")

# Plot with log scale for size
ggplot(metabolic_rates_df, aes(x = Size, y = Metabolic_Rate, color = as.factor(Species))) +
  geom_line() +
  scale_x_log10() +
  labs(title = "Metabolic Rate per Size for Each Species",
       x = "Size (w)",
       y = "Metabolic Rate",
       color = "Species") +
  theme_minimal()
```

### 1.b incorporate abundances (n_initial)

```{r}
# Function to calculate respired carbon per species and size
respired_carbon_per_species <- function(params_v1, 
                                        n = params_v1@initial_n, 
                                        n_other = params_v1@initial_n_other) {
    # Get metabolic rates per species and size (g m-2 y-1)
    metabolic_rates <- metab(params_v1)
    
    # Multiply metabolic rates by abundance for each species and size
    respired_carbon <- sweep(metabolic_rates * n, 2, params_v1@dw, "*", check.margin = FALSE)
    
    # Return the respired carbon per species and size
    return(respired_carbon)
}

# Call the function to get respired carbon
respired_carbon <- respired_carbon_per_species(params_v1)

species_names <- rownames(params_v1@species_params)
size_bins <- params_v1@w

# Extract the respiration matrix
#respiration_matrix <- respired_carbon$respired_carbon

# Assign the correct column names (size bins)
colnames(respired_carbon) <- size_bins

# Convert to a data frame and add species names
respiration_df <- as.data.frame(respired_carbon)
respiration_df$Species <- rownames(respired_carbon)

# Reshape the data to long format

respiration_long <- respiration_df %>%
  pivot_longer(cols = -Species, names_to = "Size", values_to = "respiration_production") %>%
  mutate(Size = as.numeric(Size))

# View the final data frame
print(respiration_long)

# Plot using ggplot2 with log scale for size
RC_p <- ggplot(respiration_long, aes(x = Size, y = respiration_production, color = Species)) +
  geom_line() +
  scale_x_log10() +
  labs(title = "Respired weight (g) per Size for Each Species",
       x = "Size (g)",
       y = "g C m-2 h-1",
       color = "Species") +
  theme_minimal()


```


## 1c. Convert Mizer metabolic rate (g y-1) to respiratory oxygen uptake (L h-1) then respired carbon (g C m-2 h-1)


```{r}

#RC = RO * RQ * (12/22.4) #Kwong et al. 2020
    #RC: respiratory carbon (g C ind-1 h-1); 
    #RO: respiratory oxygen uptake (L O2 ind-1 h-1); 
    #RQ: respiratory quotient (unit-less);
    #12: molar weight of carbon (g mol-1); 
    #22.4: molar volume of an ideal gas at standard pressure (mol L-1)
    #TD = time at depth

#to convert metabolic rate (respiration_production) to RO - must convert to energy (KJ) using these assumptions:

# Constants ----- include reference
energy_content <- 22         # kJ per gram based on a mixed diet - could make this more dynamic to reflect feeding types
oxygen_conversion <- 20.1      # kJ per liter of O2
hours_per_year <- 8760         # Number of hours in a year


respired_carbon_df <- respiration_long %>% 
  mutate(oxygen_uptake_liters_h = respiration_production * energy_content/oxygen_conversion/hours_per_year)

# Define the conversion factor from oxygen uptake to carbon
oxygen_to_carbon <- 12 / 22.4  # This is the factor (12/22.4) to convert oxygen to carbon

# Add RQ values to the respired_carbon_df
respired_carbon_df$RQ <- params_v1@species_params$RQ[match(respired_carbon_df$Species, params_v1@species_params$species)]

# View the updated dataframe
head(respired_carbon_df)


respired_carbon_df <- respired_carbon_df %>% 
  mutate(respired_carbon = oxygen_uptake_liters_h * RQ * oxygen_to_carbon)




```



#### plot

```{r}
# Plot using ggplot2 with log scale for size
RC_p <- ggplot(respired_carbon_df, aes(x = Size, y = respired_carbon, color = Species)) +
  geom_line() +
  scale_x_log10() +
  labs(title = "Respired Carbon per Size for Each Species",
       x = "Size (g)",
       y = "g C m-2 h-1",
       color = "Species") +
  theme_minimal()

# Display the plot
print(RC_p)

#ggsave("RC.png", plot = RC_p)

```


# 2. Fecal Carbon (POC)

## 2a. Extract feces production per species and size

```{r}

detritus_production_per_species <- function(params_v1, n = params_v1@initial_n, # in MizerShelf, detritus production includes feces, decomposing carrion and external. For now we only want to extract feces, but we'll keep the other functions in at this stage.
                                            n_other = params_v1@initial_n_other, #resource?
                                            rates = getRates(params_v1)) {
    # Calculate the consumption for each species
    consumption <- sweep((1 - rates$feeding_level) * rates$encounter * n, 2,
                         params_v1@dw, "*", check.margin = FALSE)
    
    # Calculate the feces production for each species
    feces <- sweep(consumption, 1, (1 - params_v1@species_params$alpha), "*", 
                   check.margin = FALSE)
    
    # Calculate the carrion decomposition
    carrion <- params_v1@other_params$carrion$decompose * n_other$carrion
    
    # Create a list to store detritus production per species
    detritus_per_species <- list(
        external = params_v1@other_params$detritus$external,
        feces = feces,
        carrion = carrion
    )
    
    return(detritus_per_species)
}



detritus <- detritus_production_per_species(params_v1)

species_names <- rownames(params_v1@species_params)
size_bins <- params_v1@w

# Extract the feces matrix
feces_matrix <- detritus$feces

# Assign the correct column names (size bins)
colnames(feces_matrix) <- size_bins

# Convert to a data frame and add species names
feces_df <- as.data.frame(feces_matrix)
feces_df$Species <- rownames(feces_matrix)

# Reshape the data to long format

feces_long <- feces_df %>%
  pivot_longer(cols = -Species, names_to = "Size", values_to = "Feces_production") %>%
  mutate(Size = as.numeric(Size))

# View the final data frame
print(feces_long)



```



## 2b. Convert feces (g m-2 y-1) to (g m-2 C h-1)


```{r}

# Extract the p_FC (carbon % per fecal pellet) parameter from params_v1
p_FC <- params_v1@species_params$p_FC

# Create a data frame with species and their corresponding p_FC values
p_FC_df <- data.frame(Species = rownames(params_v1@species_params), p_FC = p_FC)

# Merge the p_FC values into the feces_long dataframe
feces_long <- merge(feces_long, p_FC_df, by = "Species", all.x = TRUE)


# Calculate Fecal Carbon (FC)
feces_df <- feces_long %>%
  mutate(Fecal_Carbon = Feces_production *p_FC/hours_per_year,
          Fecal_Carbon_mg = Fecal_Carbon * 1000)  # Convert to milligrams)  


feces_p <- ggplot(feces_df, aes(x = Size, y = Fecal_Carbon_mg, color = as.factor(Species))) +
  geom_line() +
  scale_y_log10()+
  scale_x_log10() +
  labs(title = "Fecal Carbon per Size for Each Species",
       x = "Size (g)",
       y = "mg C m-2 h-1",
       color = "Species") +
  theme_minimal()

feces_p

#ggsave("FC.png", plot = feces_p,  width = 7, height = 5)

```



# 3. Carrion carbon (POC)

## 3a. Approach 1: take carrion from proportion of detritus. However, this includes a degradation term, so we may miss "fresh" carrion. Also, if used with degradation term we may not need to apply remineralization rates later... i.e. this "POC" is actually "DIC" and remains suspended.

```{r}
#approach 1 - take carrion from proportion of detritus

# # Extract the feces matrix
# carrion_matrix <- detritus$carrion
# 
# # Assign the correct column names (size bins)
# colnames(carrion_matrix) <- size_bins
# 
# # Convert to a data frame and add species names
# carrion_df <- as.data.frame(carrion_matrix)
# carrion_df$Species <- rownames(carrion_matrix)
# 
# # Reshape the data to long format
# 
# carrion_long <- carrion_df %>%
#   pivot_longer(cols = -Species, names_to = "Size", values_to = "carrion_production") %>%
#   mutate(Size = as.numeric(Size))
# 
# # View the final data frame
# print(carrion_long)



```

## 3b. Approach 2: manually extract species specific external mortality rates and multiply by weight. However, extmort is constant so size is not taken into account.


```{r}
# Function to calculate respired carbon per species and size
carcass_carbon_per_species <- function(params_v1, 
                                        n = params_v1@initial_n, 
                                        n_other = params_v1@initial_n_other) {
    # Get metabolic rates per species and size (g m-2 y-1)
    ext_mort_rates <- getExtMort(params_v1)
    
    # Multiply metabolic rates by abundance for each species and size
    ext_mort_rates <- sweep(ext_mort_rates * n, 2, params_v1@dw, "*", check.margin = FALSE)
    
    # Return the respired carbon per species and size
    return(ext_mort_rates)
}

# Call the function to get respired carbon
ext_mort_rates <- carcass_carbon_per_species(params_v1)

species_names <- rownames(params_v1@species_params)
size_bins <- params_v1@w



# Assign the correct column names (size bins)
colnames(ext_mort_rates) <- size_bins

# Convert to a data frame and add species names
ext_mort_rates_df <- as.data.frame(ext_mort_rates)
ext_mort_rates_df$Species <- rownames(ext_mort_rates)

# Reshape the data to long format

ext_mort_rates_long <- ext_mort_rates_df %>%
  pivot_longer(cols = -Species, names_to = "Size", values_to = "ext_mort_rates") %>%
  mutate(Size = as.numeric(Size))

# View the final data frame
print(ext_mort_rates_long)

# Plot using ggplot2 with log scale for size
CC_p <- ggplot(ext_mort_rates_long, aes(x = Size, y = ext_mort_rates, color = Species)) +
  geom_line() +
  scale_x_log10() +
  labs(title = "Carcass weight (g) per Size for Each Species",
       x = "Size (g)",
       y = "g C m-2 y-1",
       color = "Species") +
  theme_minimal()

```


```{r}


carcass_carbon_df <- ext_mort_rates_long %>% 
  mutate(g_C = ext_mort_rates * 0.4) %>% # change this to species specific carbon content (or leave as assumption - Steinberg et al. 2017?)
  mutate(carcass_C = g_C * 0.1713287/hours_per_year) #constant taken from MizerShelf - proportion of ext_mort that produces carrion (i.e. some of the external mortality will not contribute to carrion production due to consumption from non-modeled species like sea birds)


carrion_p <- ggplot(carcass_carbon_df, aes(x = Size, y = carcass_C, color = as.factor(Species))) +
  geom_line() +
  scale_y_log10()+
  scale_x_log10() +
  labs(title = "Carcass Carbon per Size for Each Species",
       x = "Size (g)",
       y = "g C m-2 h-1",
       color = "Species") +
  theme_minimal()

carrion_p

#ggsave("CC.png", plot = carrion_p)

```



## 3c. Extract external contribution to carrion production (MizerShelf). Ideal approach, but unable to partition natural mortality atm.

```{r}

# #approach 3: extract external contribution to carrion production
# 
# carrion_production_per_species <- function(params_v1, n = params_v1@initial_n,
#                                            rates = getRates(params_v1)) {
#     # Calculate natural mortality contribution to carrion per species
#     mu_b_contribution <- (params_v1@mu_b * n) * (params_v1@w * params_v1@dw) *
#                          params_v1@other_params$carrion$ext_prop
# 
#     # Calculate gear mortality contribution to carrion per species
#     gear_mort_contribution <- (gearMort(params_v1, rates$f_mort) * n) *
#                               (params_v1@w * params_v1@dw)
# 
#     # Calculate discards contribution to carrion per species
#     discards_contribution <- ((rates$f_mort * n) * (params_v1@w * params_v1@dw)) *
#                              params_v1@species_params$discard
# 
#     # Create a list to store carrion production per species
#     carrion_per_species <- list(
#         mu_b = mu_b_contribution,
#         gear_mort = gear_mort_contribution,
#         discards = discards_contribution
#     )
# 
#     return(carrion_per_species)
# }
# 
# # Assume params is already defined and get carrion production per species
# carrion <- carrion_production_per_species(params_v1)
# 
# carrion
# 
# getCarrionProduction(params_v1)


```


# 4. Carbon pools

## 4a. combine carbon products dataframe


```{r}

# First, merge respired_carbon_long and feces_df
combined_df1 <- merge(respired_carbon_df, feces_df, by = c("Species", "Size"))

# Then, merge the result with carcass_carbon_df
combined_df <- merge(combined_df1, carcass_carbon_df, by = c("Species", "Size"))

# View the combined dataframe
head(combined_df)

```

## 4b. add in DVM info to split carbon products up by depth layer - based on vertical_migration_array

```{r}

# Convert vertical_migration_array to a dataframe
vertical_migration_df <- as.data.frame(vertical_migration_array)

# Transpose the dataframe so species names are in a column
vertical_migration_df <- t(vertical_migration_df)

# Convert it back to a dataframe and add row names as a column
vertical_migration_df <- as.data.frame(vertical_migration_df)
vertical_migration_df$Species <- rownames(vertical_migration_df)
rownames(vertical_migration_df) <- NULL


# Rename the columns for easier merging
colnames(vertical_migration_df) <- c("P_Epi", "P_Umeso", "P_Lmeso", "Species")


# Split the Species column into species name and size using strsplit
split_values <- strsplit(as.character(vertical_migration_df$Species), "(?<=\\D)\\.(?=\\d)", perl = TRUE)

# Extract the species name
vertical_migration_df$Species <- sapply(split_values, `[`, 1)

# Extract the size
vertical_migration_df$Size <- sapply(split_values, `[`, 2)
vertical_migration_df <- vertical_migration_df %>% 
  select(-Size)


# Merge the vertical migration data with your combined dataframe
final_df <- merge(combined_df, vertical_migration_df, by = c("Species"), all.x = TRUE) #may need to make this more dynamic and incorporate DVM based on size - to be discussed later

final_df <- final_df %>% 
  unique()
# View the final dataframe
head(final_df)

```



## 4c. Create Carbon pools dataframe

```{r}

DIC_POC <- final_df %>% 
rowwise() %>%
  mutate(total_DIC = sum(respired_carbon),
         total_POC = sum(Fecal_Carbon, carcass_C)) %>%
  ungroup()


carbon_pools <- DIC_POC %>% 
  select(Species, total_DIC, total_POC, P_Epi, P_Umeso, P_Lmeso) %>% 
  group_by(Species) %>% 
  mutate(g_POC_m2_h = sum(total_POC),
         g_DIC_m2_h = sum(total_DIC)) %>% 
  ungroup() %>% 
  select(Species, g_POC_m2_h, g_DIC_m2_h,P_Epi, P_Umeso, P_Lmeso) %>% 
  unique() %>% 
    mutate(DIC_Epi = g_DIC_m2_h *15.8* P_Epi, #based on % of species specific biomass per layer and multiplied by daylight at SOTS during SOLACE (15.8 h)
         DIC_Umeso = g_DIC_m2_h *15.8* P_Umeso,
         DIC_Lmeso = g_DIC_m2_h *15.8* P_Lmeso,
         POC_Epi = g_POC_m2_h *15.8* P_Epi,
         POC_Umeso = g_POC_m2_h *15.8* P_Umeso,
         POC_Lmeso = g_POC_m2_h *15.8* P_Lmeso)



```

###Group by higher taxonomic level

```{r}


# Add the higher_taxa column based on Species
carbon_pools <- carbon_pools %>%
  mutate(higher_taxa = case_when(
    Species == "Atolla wyvillei" ~ "Gelatinous (other)",
    Species == "Gymnosomata" ~ "Gelatinous (other)",
    Species == "Chaetognath" ~ "Gelatinous (other)",
    Species == "Cranchidae" ~ "Cranchidae",
    Species == "Pyrosoma atlanticum" ~ "P. atlanticum",
    Species == "Euphausia similis" ~ "Krill",
    Species == "Thysannoessa macura" ~ "Krill",
    Species == "Bathylagus spp" ~ "Fish",
    Species == "Cyclothone spp" ~ "Fish",
    Species == "Lampanyctus australis" ~ "Fish",
    Species == "Nannobrachium achirus" ~ "Fish",
    Species == "Sergestes spp" ~ "Crustacean (other)",
    Species == "Themisto spp" ~ "Crustacean (other)",
    TRUE ~ NA_character_ # Assign NA for any unmatched species
  ))

# View the updated dataframe
print(carbon_pools)

```


## 4d. Plot respired DIC 

```{r}

# Gather the data into a long format for plotting
DIC_injection_resp_df <- carbon_pools %>%
  select(Species, DIC_Epi, DIC_Umeso, DIC_Lmeso, higher_taxa) %>%
  pivot_longer(cols = c(DIC_Epi, DIC_Umeso, DIC_Lmeso), 
               names_to = "Layer", 
               values_to = "DIC") %>%
  mutate(
    Layer = recode(Layer,
                   "DIC_Epi" = "Epi",
                   "DIC_Umeso" = "Umeso",
                   "DIC_Lmeso" = "Lmeso"),
    Layer = factor(Layer, levels = c("Lmeso", "Umeso", "Epi")),
    total_DIC_mg = DIC * 1000
  )


# Create the horizontal bar plot
DIC_injection_resp_p <- ggplot(DIC_injection_resp_df, aes(x = total_DIC_mg, y = Layer, fill = Species)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "mg C m-2 d-1", y = "Depth Stratum") +
  xlim(0,10)+
  scale_y_discrete(labels = c("Lmeso","Umeso","Epi")) +
  theme(legend.title = element_blank())

DIC_injection_resp_p

ggsave("Supp_fig8_DICpertaxa.png", plot = DIC_injection_resp_p, width = 7, height = 5)
```


```{r}
DIC_injection_total <- DIC_injection_resp_df %>% 
  group_by(Layer) %>% 
  mutate(mg_C_m2_d_layer = sum(total_DIC_mg),
         POC_DIC = "DIC_inject") %>% 
  ungroup() %>% 
  select(Layer, mg_C_m2_d_layer, POC_DIC) %>% 
  unique() 
```



## 4e. Plot POC (fecal & carcass carbon)

```{r}


# Gather the data into a long format for plotting
POC_injection_df <- carbon_pools %>%
  select(Species, POC_Epi, POC_Umeso, POC_Lmeso, higher_taxa) %>%
  pivot_longer(cols = c(POC_Epi, POC_Umeso, POC_Lmeso), 
               names_to = "Layer", 
               values_to = "POC") %>%
  mutate(Layer = factor(Layer, levels = c("POC_Lmeso", "POC_Umeso", "POC_Epi")),
    Layer = recode(Layer,
                   "POC_Epi" = "Epi",
                   "POC_Umeso" = "Umeso",
                   "POC_Lmeso" = "Lmeso"),
    Layer = factor(Layer, levels = c("Lmeso", "Umeso", "Epi")),
    total_POC_mg = POC * 1000
  )

# Create the horizontal bar plot
POC_injection_p <- ggplot(POC_injection_df , aes(x = total_POC_mg, y = Layer, fill = Species)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "mg C m-2 d-1", y = "Depth Stratum") +
  xlim(0,10)+
  scale_y_discrete(labels = c("Lmeso","Umeso","Epi")) +
  theme(legend.title = element_blank())

POC_injection_p

ggsave("Supp_fig9_POCpertaxa.png", plot = POC_injection_p, width = 7, height = 5)


```

#plot higher_taxa carbon pools (Figure 3)

```{r}

#colors (to match Baker et al., 2025)

micronekton_colors <- plasma(6, begin = 0, end = 0.8)   


# Combine micronekton and mesozooplankton colors
taxa_colors <- c(
  "Gelatinous (other)" = micronekton_colors[1],
  "P. atlanticum" = micronekton_colors[2],
  "Krill" = micronekton_colors[3],
  "Cranchidae" = micronekton_colors[4],
  "Crustacean (other)" = micronekton_colors[5],
  "Fish" = micronekton_colors[6]
)

#DIC

# Create the horizontal bar plot
DIC_injection_highertaxa_p <- ggplot(DIC_injection_resp_df, aes(x = total_DIC_mg, y = Layer, fill = higher_taxa)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "mg C m-2 d-1", y = "Depth Stratum") +
  xlim(0,10)+
  scale_y_discrete(labels = c("Lmeso","Umeso","Epi")) +
  scale_fill_manual(values = taxa_colors, name = NULL)+
  theme(legend.title = element_blank())

DIC_injection_highertaxa_p

```

```{r}
#POC


# Create the horizontal bar plot
POC_injection_highertaxa_p <- ggplot(POC_injection_df , aes(x = total_POC_mg, y = Layer, fill = higher_taxa)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "mg C m-2 d-1", y = "Depth Stratum") +
  xlim(0,10)+
  scale_y_discrete(labels = c("Lmeso","Umeso","Epi")) +
  scale_fill_manual(values = taxa_colors, name = NULL)+
  theme(legend.title = element_blank())

POC_injection_highertaxa_p

```


```{r}
# Combine the two plots side by side with a shared legend below
combined_plot <- DIC_injection_highertaxa_p + POC_injection_highertaxa_p +
  plot_layout(guides = "collect", ncol = 2) + # Arrange plots in 2 columns and collect the legend
  plot_annotation(tag_levels = 'A') & # Add tags for panels
  theme(legend.position = "bottom") # Place the legend below

# Display the combined plot
combined_plot

ggsave("Fig3_DIC_POC.png", plot = combined_plot, width = 7, height = 5)

```

##### this is incorrect............................


## 4f. add in microbial remineralisation rates

```{r}

#From Petiteau et al. submitted: average microbial remin rate (Kmicrob) from C-RESPIRE trap = 0.11 +/- 0.05 d-1
# attempt 1 - apply this rate to POC through entire water column

# # microbial remineralization rate 
# Kmicrob <- 0.11  #Petiteau et al. SOTS during SOLACE

convert_POC_to_DIC <- function(POC_injection_df, Kmicrob = 0.11) {
  
  # Create an empty dataframe to store DIC results
  DIC_long <- POC_injection_df
  DIC_long$DIC <- 0  # Initialize DIC column with zeros

  # Get unique species
  species_list <- unique(POC_injection_df$Species)

  # Iterate over each species
  for (species in species_list) {
    
    # Subset POC data for the species
    POC_species <- subset(POC_injection_df, Species == species)
    
    # Calculate DIC for the Epi layer
    DIC_Epi <- POC_species$POC[POC_species$Layer == "Epi"] * Kmicrob
    
    # Calculate POC that sinks to Umeso
    POC_Umeso_influx <- POC_species$POC[POC_species$Layer == "Epi"] * (1 - Kmicrob)
    
    # Calculate DIC for the Umeso layer
    DIC_Umeso <- (POC_species$POC[POC_species$Layer == "Umeso"] + POC_Umeso_influx) * Kmicrob
    
    # Calculate POC that sinks to Lmeso
    POC_Lmeso_influx <- (POC_species$POC[POC_species$Layer == "Umeso"] + POC_Umeso_influx) * (1 - Kmicrob)
    
    # Calculate DIC for the Lmeso layer
    DIC_Lmeso <- (POC_species$POC[POC_species$Layer == "Lmeso"] + POC_Lmeso_influx) * Kmicrob
    
    # Update the DIC_long dataframe with the new DIC values
    DIC_long$DIC[DIC_long$Species == species & DIC_long$Layer == "Epi"] <- DIC_Epi
    DIC_long$DIC[DIC_long$Species == species & DIC_long$Layer == "Umeso"] <- DIC_Umeso
    DIC_long$DIC[DIC_long$Species == species & DIC_long$Layer == "Lmeso"] <- DIC_Lmeso
  }

  return(DIC_long)
}


DIC_to_POC_df <- convert_POC_to_DIC(POC_injection_df)

DIC_to_POC_df<- DIC_to_POC_df %>% 
  select(-POC) %>% 
  mutate(total_DIC_mg = DIC * 1000)

# Create the horizontal bar plot
POC_to_DIC_p <- ggplot(DIC_to_POC_df, aes(x = total_DIC_mg, y = Layer, fill = Species)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "DIC (mg C m-2 d-1)", y = "Depth Stratum") +
  scale_y_discrete(labels = c("Lmeso","Umeso","Epi")) +
  theme(legend.title = element_blank())

# Plot the graph
POC_to_DIC_p

#ggsave("POC_to_DIC_p.png", plot = POC_to_DIC_p, width = 7, height = 5)

```

#plot POC injection and DIC accumulation after remineralization

```{r}
POC_injection_layer <- POC_injection_df %>% 
  group_by(Layer) %>% 
  mutate(mg_C_m2_d_layer = sum(total_POC_mg)) %>% 
  ungroup() %>% 
  mutate(POC_DIC = "POC_inject") %>% 
  select(Layer, mg_C_m2_d_layer, POC_DIC) %>% 
  unique()


POC_to_DIC_layer <- DIC_to_POC_df %>% 
  group_by(Layer) %>% 
   mutate(mg_C_m2_d_layer = sum(total_DIC_mg)) %>% 
  ungroup() %>% 
  mutate(POC_DIC = "POC_rem_DIC") %>% 
  select(Layer, mg_C_m2_d_layer, POC_DIC) %>% 
  unique()

# Merge DIC_long and POC_DIC_long by Species and layer
POC_DIC_df <- rbind(POC_injection_layer, POC_to_DIC_layer)

```
## How much POC sinks below meso?

```{r}
#Add how much POC was injected over whole water column and subtract the total DIC remineralized to determine how much sinks below meso

POC_to_bathy <- POC_DIC_df %>% 
  group_by(POC_DIC) %>% 
  mutate(total = sum(mg_C_m2_d_layer)) %>% 
  ungroup() %>% 
  select(POC_DIC, total) %>% 
  unique() 


# Extract the numeric values using the 'total' column
POC_inject <- POC_to_bathy$total[POC_to_bathy$POC_DIC == "POC_inject"]
POC_rem_DIC <- POC_to_bathy$total[POC_to_bathy$POC_DIC == "POC_rem_DIC"]

# Calculate the difference
DIC_bathy <- POC_inject - POC_rem_DIC

# Print the result
print(DIC_bathy)


# 10.13 mg C m-2 d-1 sinks below 1000 m

```


## 4g. combine with respired DIC (note - plot not included in ms)

```{r}

# Generate 13 colors spaced along the palette
#taxa_colors <- viridis(13, begin = 0, end = 0.9)
taxa_colors <- plasma(13, begin = 0, end = 0.9)   


# Combine micronekton and mesozooplankton colors
taxa_colors <- c(
  "Pyrosoma atlanticum" = taxa_colors[1],
  "Gymnosomata" = taxa_colors[2],
  "Themisto spp" = taxa_colors[3],
  "Thysannoessa macura" = taxa_colors[4],
  "Euphausia similis"= taxa_colors[5],
  "Cyclothone spp" = taxa_colors[6],
  "Bathylagus spp" = taxa_colors[7],
  "Lampanyctus australis" = taxa_colors[8],
  "Nannobranchium achirus" = taxa_colors[9],
  "Sergestes spp" = taxa_colors[10],
  "Atolla spp" = taxa_colors[11],
  "Cranchidae" = taxa_colors[12],
  "Chaetognath" = taxa_colors[13]
)

# Define the desired order for `higher_taxa`
desired_order <- c(
  "P. atlanticum",
  "Gymnosomata",
  "Themisto spp",
  "Thysannoessa macura",
  "Euphausia similis",
  "Cyclothone spp",
  "Bathylagus spp",
  "Lampanyctus australis",
  "Nannobranchium achirus",
  "Sergestes spp",
  "Atolla spp",
  "Cranchidae",
  "Chaetognath"
)

# View the colors
taxa_colors



# Merge DIC_long and POC_DIC_long by Species and layer
merged_df <- left_join(DIC_to_POC_df, DIC_injection_resp_df, by = c("Species", "Layer"))

merged_df <- merged_df %>%
  mutate(
    total_DIC_mg = total_DIC_mg.x +total_DIC_mg.y) %>%   # Assuming DIC from DIC_to_POC_df is DIC.x and from POC_DIC_long is DIC.y
  #total_DIC_mg = Total_DIC * 1000) %>%   # Convert to milligrams)  
  select(Species, Layer, total_DIC_mg)
  #factor(
  #merged_df$Species
  #levels = desired_order
#)


# Create the horizontal bar plot
Total_DIC_pools <- ggplot(merged_df, aes(x = total_DIC_mg , y = Layer, fill = Species)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "DIC (mg C m-2 d-1)", y = "Depth Stratum") +
  xlim(0,10)+
  scale_y_discrete(labels = c("Lmeso","Umeso","Epi")) +
  scale_fill_manual(values = taxa_colors, name = NULL)+
  theme(legend.title = element_blank()) +
  theme(
    axis.text = element_text(size = 14),  # Adjust size of axis text
    axis.title = element_text(size = 16), # Adjust size of axis titles
    legend.text = element_text(size = 12), # Adjust size of legend text
    legend.title = element_text(size = 14) # Adjust size of legend title
  )


# Plot the graph
Total_DIC_pools




```




```{r}
DIC_per_layer <- merged_df %>% 
  group_by(Layer) %>% 
  mutate(total_layer_DIC_mg = sum(total_DIC_mg)) %>% 
  ungroup() %>% 
  group_by(Species, Layer) %>% 
  mutate(percent_cont = total_DIC_mg/total_layer_DIC_mg) %>% 
  ungroup()

DIC_final_pools <- DIC_per_layer %>% 
  select(total_layer_DIC_mg, Layer) %>% 
  unique() %>% 
  rename(mg_C_m2_d_layer = total_layer_DIC_mg) %>% 
  mutate(POC_DIC = "Total DIC")

#add DIC_bathy

# Add a new row with mutate and bind_rows
DIC_final_pools <- DIC_final_pools %>% 
  bind_rows(
    tibble(
      mg_C_m2_d_layer = DIC_bathy,
      Layer = "Bathy",
      POC_DIC = "Total DIC"
    )
  )

# View the updated dataframe
print(DIC_final_pools)


```
#plot carbon pools

```{r}

# Ensure the correct stacking order
DIC_final_pools$Layer <- factor(DIC_final_pools$Layer, levels = c("Epi", "Umeso", "Lmeso", "Bathy"))

# Create a custom blue gradient by replacing the darkest shade
custom_blue_palette <- function(n) {
  pal <- viridis_pal(option = "mako")(n) # Generate the mako palette
  pal[1] <- "#00008B" # Replace the darkest color with a dark blue (hex: DarkBlue)
  return(rev(pal))
}

# Create the stacked bar chart with renamed legend labels
total_DIC_WC <- ggplot(DIC_final_pools, aes(x = "", y = mg_C_m2_d_layer, fill = Layer)) +
  geom_bar(stat = "identity", width = 0.6) +
  scale_fill_manual(
    values = custom_blue_palette(4), # Use the custom blue palette
    labels = c(
      "Epi" = "Epi 0 - 200 m",
      "Umeso" = "Umeso 200 - 400 m",
      "Lmeso" = "Lmeso 400 - 1000 m",
      "Bathy" = "Bathy >1000 m"
    )
  ) +
  labs(
    title = NULL,
    x = NULL,
    y = "mg C m² d⁻¹",
    fill = "Layer"
  ) +
  theme_minimal() +
  theme(
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()
  )

# Save the plot
ggsave("Fig4_DIC_WC.png", plot = total_DIC_WC, width = 7, height = 5)


```



```{r}
carbon_pools_layer <- rbind(POC_DIC_df, DIC_injection_total, DIC_final_pools)

carbon_pools_layer_pivoted <- carbon_pools_layer %>%
  pivot_wider(
    names_from = POC_DIC,            # Values in 'POC_DIC' become column names
    values_from = mg_C_m2_d_layer   # Values from 'mg_C_m2_d_layer' fill the new columns
  )




write.csv(carbon_pools_layer_pivoted, "Table3_Total_carbon_pools.csv")

```



#Compare values with chapter 1: remove species where % epi = 0 - also need to remove POC sinking contribution.... 

```{r}
#POC_injection_df -= POC with no remineralisation
#DIC_injection_resp_df = respired DIC


# Merge DIC_long and POC_DIC_long by Species and layer
merged_df_no_remin <- left_join(POC_injection_df, DIC_injection_resp_df, by = c("Species", "Layer"))

merged_df_no_remin <- merged_df_no_remin %>%
  mutate(
    Total_DIC = POC + DIC,  # Assuming DIC from DIC_to_POC_df is DIC.x and from POC_DIC_long is DIC.y
  total_DIC_mg = Total_DIC * 1000) %>%   # Convert to milligrams)  
  select(Species, Layer, total_DIC_mg )


merged_df_no_remin <- merged_df_no_remin %>% 
  mutate(mod_type = "modeled")



# N. achirus, Cyclothone spp, Bathylagus spp, Gymnosomata, atolla wyvillei

df_no_epi <- merged_df_no_remin  %>% 
  filter(!Species %in% c("Nannobrachium achirus", "Cyclothone spp", "Gymnosomata", "Bathylagus spp", "Atolla wyvillei"))

write.csv(df_no_epi, "modeled_DIC_no_epi.csv")

# Create the horizontal bar plot
Total_DIC_pools_no_remin <- ggplot(df_no_epi, aes(x = total_DIC_mg , y = Layer, fill = Species)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "DIC (mg C m-2 d-1)", y = "Depth Stratum") +
  scale_y_discrete(labels = c("Lmeso","Umeso","Epi")) +
  theme(legend.title = element_blank()) +
  theme(
    axis.text = element_text(size = 14),  # Adjust size of axis text
    axis.title = element_text(size = 16), # Adjust size of axis titles
    legend.text = element_text(size = 12), # Adjust size of legend text
    legend.title = element_text(size = 14) # Adjust size of legend title
  )


# Plot the graph
Total_DIC_pools_no_remin

```





